---
title: "Scottish agriculture greenhouse gas and nitrogen use: 2022-23"
author: no
date: "27 June 2024"
output:
  html_document:
    template: _template.html
    toc: yes
    toc_float: yes
    toc_depth: 2
    self_contained: yes
sgtemplates:
  footer:
    link:
      - href: "accessibility.html"
        text: "Accessibility"
      - href: "contact.html"
        text: "Contact"
  header:
    site_branding: "Scottish agriculture greenhouse gas and nitrogen use: 2022-23"
    toc:
      sticky: false
  seo:
    description: "SEO (search engine optimisation) description"
    keywords: "agriculture 2022, agriculture emissions, scotland ghg, beef emissions"
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
source("scripts/national_charts.R")
source("scripts/CN_analysis_build.R")

```

This report includes:

-   Estimates of greenhouse gas emissions for Scottish agriculture as reported in the [Scottish Greenhouse Gas (GHG) Statistics 2022](https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022/)

-   Estimates of greenhouse gas emissions by agricultural subsector, based on [analysis by Scotland's Rural College (SRUC)](https://www.gov.scot/publications/disaggregating-headline-smart-inventory-figures/).

-   Subsector estimates break-down total greenhouse gas emissions by subsector rather than by source. This differs from the presentation in the Scottish Greenhouse Gas Statistics, where emissions sources are not fully allocated to agricultural subsectors. For example, mobile machinery emissions are allocated to each subsector in this report, whilst they were previously reported in total.

-   Emission and nitrogen use estimates for the average farm based on commercial sized farms in the Farm Business Survey. Farms are complex businesses with multiple activities that contribute to GHG emissions. For example, cereal farms may have livestock and the scale of this secondary activity can vary.

# Key Points

-   In 2022, greenhouse gas emissions from Scottish agriculture were at their lowest levels since the baseline period in 1990. Total emissions from agriculture in 2022 were 7.7 million tonnes of carbon dioxide equivalent (MtCO~2~e). This is a decrease of 3% (0.3 MtCO~2~e) compared with 2021.

-   Falling emissions from arable~~,~~ dairy and suckler beef subsectors drove the year-on-year decrease in greenhouse gas emissions from agriculture in 2022. Arable and suckler beef emissions both fell by 0.1 MtCO~2~e from the previous year.

-   The suckler beef subsector is the largest contributor to net agriculture emissions. In 2022 emissions were 3.2 MtCO~2~e, 41% of agriculture emissions. Suckler beef emissions fell by 0.6 MtCO~2~e from 1990 to 2022 - the largest absolute reduction across the agriculture subsectors. This follows the trend of declining cattle numbers over the same period.

-   Emissions from arable farming increased by 14% from 1990 to 2022. Arable farming accounts for the majority of emissions from liming and fuel combustion.

# Introduction

## What you need to know

This report provides greenhouse gas (GHG) emissions and nitrogen use estimates for agriculture at sectoral level, subsectoral level, and for average Scottish farms by farm type. Methodology differences mean average farm estimates are not directly comparable with national estimates.

It is important to note that agriculture and food production rely on natural processes that will always cause a degree of greenhouse gas emissions.

## About greenhouse gas emissions in agriculture

There are three main greenhouse gases produced in agriculture. These are carbon dioxide, methane, and nitrous oxide. Total emissions are presented in carbon dioxide equivalent (CO~2~e) units. This takes into account the different effects that the different gases have on climate change, known as their global warming potential (GWP). Over a 100-year period, methane's GWP is considered to be 28 times stronger than that of carbon dioxide, while nitrous oxide's GWP is considered to be 265 times stronger than carbon dioxide.

## About the national greenhouse gas estimates for agriculture and agricultural subsectors

The data source for the national estimates is the [Scottish Greenhouse Gas Statistics 2022](https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022/). Emissions are allocated by sector and type of gas.

Subsector analysis breaks down each sector into enterprise or activity type. This gives more detail about the emission sources within each sector. For Agriculture, the X reported in the Scottish Greenhouse Gas Statistics are:

-   Agricultural combustion

-   Agricultural soils

-   Livestock

-   Other agriculture -- including lubricants and crop combustion

This report provides emission estimates for the following subsectors:

-   suckler beef

-   dairy beef

-   dairy

-   sheep

-   arable

-   other -- including pigs, other livestock and non-agriculture.

The emission sources are allocated to agricultural subsectors in proportions that are given in [table 1]. Proportions are based on [research by SRUC](https://www.gov.scot/publications/disaggregating-headline-smart-inventory-figures/) from 2021 and have been applied to the entire timeseries.

#### Table 1

Further information about the data sources and methods are avaliable in the Methodology [LINK].

## About nitrogen in agriculture

Nitrogen is an essential nutrient for plant growth. Nitrogen supplied by fertiliser, soil and other sources are consumed in harvested crops and animal products. However, excess nitrogen can also be lost to the wider environment. Losses of nitrogen to the environment can have harmful effects on, for example, climate change, biodiversity, air quality and water quality.

## About national nitrogen use estimates

National results for agriculture nitrogen use efficiency are available in the [Scottish Nitrogen Balance Sheet](https://www.gov.scot/publications/scottish-nitrogen-balance-sheet-2021/).

Note that the data are for 2021 and are not comparable with the whole-farm average nitrogen use estimates in this report.

## About the average whole-farm GHG emissions and nitrogen use estimates

Average whole-farm estimates complement the national results and enterprise specific estimates by providing a view and timeseries of emissions and nitrogen use on real farms. Farms are complex businesses with multiple enterprises that contribute to GHG emissions. For example, cereal farms may have livestock and the scale of this secondary enterprise can vary.

Data do not cover the full agricultural industry and are not comparable with national estimates.

This analysis is based on a survey of around 400 commercial farms (whose standard output exceeds around £20,000 per year). The sample excludes farm types not in receipt of support payments (such as pigs, poultry and horticulture). The survey sample is weighted using results from the 2022 June Agricultural Census. Results are shown for the average (median) farm and as such are not directly comparable to results published on farm income.

On-farm gross emissions are estimated using a life cycle assessment (LCA) based carbon calculator ([Agrecalc](https://www.agrecalc.com/)). They do not account for any sequestration by woodland areas or soil. Nitrogen estimates are based on standard estimates of nitrogen content in all farm inputs and outputs where possible. Nitrogen estimates are not included for organic farms, which means a small proportion of the sample are excluded.

These data are designated as official statistics in development. They are newly developed statistics undergoing testing. More information is available in the Methodology [LINK]

## Measuring whole-farm greenhouse gas emissions

Greenhouse gas (GHG) emissions are measured by:

-   Absolute gross GHG emissions (tonnes CO~2~ equivalent per hectare)

-   Emissions intensity (kg CO~2~ equivalent per kg of output)

These indicators measure GHG emissions from farm production activities. They do not account for any sequestration by woodland areas or soil.

Absolute gross emissions measures the GHG emissions from a unit area of agriculture land (hectare).

Emissions intensity measures the carbon footprint of a unit of product up to the farm gate. It includes all types of agricultural products produced on the farm.

Data collected through the Farm Business Survey provide a consistent and high-quality input data set about farm activities. GHG emissions are then estimated using a carbon footprint calculator tool ([Agrecalc](https://www.agrecalc.com/)). It is based on the life cycle assessment (LCA) framework for estimating emissions from products and processes. The LCA accounts for emissions up to when product leaves the farm. The tool uses the latest IPCC Tier I and Tier II as well as national figures from the UK National Greenhouse Gas Inventory in its calculations. It is PAS2050:2011 certified. Different results may be obtained using different carbon calculators or different methods, for example using an inventory approach.

## Measuring whole-farm nitrogen use

Nitrogen use is measured by:

-   Nitrogen balance (kg N surplus per hectare)

-   Nitrogen use efficiency NUE (% N output / N input)

Nitrogen balance is the difference between total nitrogen input and output. A higher balance indicates less efficient use of nitrogen. It provides an estimate of the size of the nitrogen surplus not being captured in agricultural products that is potentially available for losses.

Nitrogen use efficiency (NUE) is the ratio of nitrogen outputs to inputs. It indicates the proportion of nitrogen used in the farm system. This measure allows for better comparison across farms. NUE values should always be interpreted in relation to nitrogen surpluses and nitrogen outputs.

A higher NUE typically indicates a more efficient use of nitrogen but very high values may indicate unsustainable "soil mining". The best range of NUE values depends on the type of farming activity as well as environmental conditions, livestock types and feed types.

Nitrogen estimates are based on standard estimates of nitrogen content in farm inputs and outputs. The amount of nitrogen is estimated for all farm inputs and outputs up to farm gate, where possible. Limited information is available for some inputs and outputs.

## An Official Statistics in Development Publication for Scotland

These statistics are official statistics in development. Official statistics in development may be new or existing statistics, and will be tested with users, in line with the standards of trustworthiness, quality, and value in the [Code of Practice for Statistics](https://code.statisticsauthority.gov.uk/the-code/).

Scottish Government statistics are regulated by the Office for Statistics Regulation (OSR). OSR sets the standards of trustworthiness, quality and value in the [Code of Practice for Statistics](https://code.statisticsauthority.gov.uk/the-code/) that all producers of official statistics should adhere to.

# Agriculture sector greenhouse gas emissions

Greenhouse gas emissions from Scottish agriculture were 7.7 million tonnes of carbon dioxide equivalent (MtCO~2~e) in 2022. This is the lowest value since the baseline period in 1990, a decrease of 12% or 1.0 MtCO~2~e over this period. Compared with the previous year, total agriculture emissions in 2022 fell by 3% from 8.0 MtCO~2~e in 2021.

A fall in emissions from arable, dairy and suckler beef subsectors drove the decrease in total agriculture emissions in 2022. Both arable and suckler beef emissions fell by 0.1 MtCO~2~e from the previous year. Emissions from dairy beef also fell by 0.05 MtCO~2~e.

#### Greenhouse gas emissions from agriculture at lowest levels

Total and agriculture greenhouse gas emissions in Scotland, 1990 to 2022
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig1, echo = FALSE}
fig1

```

Prior to the decline of 2022, greenhouse gas emissions from Scottish agriculture were relatively stable. From 2017 to 2021 emissions were 7.9 MtCO~2~e and rose or fell by 0.1 MtCO~2~e on average. Since 2000, agriculture emissions have remained below figures recorded in 1990 (8.8 MtCO~2~e).

More results for national GHG emissions, including for Agriculture are available in the [Scottish Greenhouse Gas Statistics](https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022).

# Agriculture sub-sector greenhouse gas emissions

The suckler beef subsector has the largest share of Scottish agriculture emissions from 1990 to 2022. However, it has also reduced total agriculture emissions by the greatest amount compared with all other subsectors. From 1990 to 2022 suckler beef emissions have fallen by 0.6 MtCO~2~e. This has been driven by the decline in the number of cattle in Scotland over the same period. Cattle numbers have fallen from 2.1 million in 1990 to 1.7 million in 2022, a decrease of 20%.

Emissions from arable farming in Scotland have increased by 14% from 1990 to 2022.

#### Figure 2: Suckler beef subsector has the largest share of Scottish agriculture greenhouse gas emissions

Agriculture greenhouse gas emissions in Scotland, by subsector, 2022

```{r fig2, echo = FALSE}
fig2


```

## Sources of agriculture greenhouse gas emissions

The possible sources of greenhouse gas emissions from agricultural activity are grouped as the following:

-   **enteric fermentation** - a digestive process whereby carbohydrates are broken down by microorganisms into simple molecules. Both ruminant animals (e.g. cattle and sheep), and some non-ruminant animals (e.g. swine and horses) produce methane, although ruminants are the largest source per unit of feed intake

-   **manure management** - methods used to store or treat manure. These methods influence emission rates of methane. Liquid manure decomposes anaerobically and releases a large amount of methane. Solid manure decomposes aerobically and produces little to no methane.

-   **agricultural soils**

-   **fuel combustion** - including emissions from static and mobile machinery

-   **liming** - a soil management technique where lime (e.g. calcium carbonate) is applied to acidic soils. Liming neutralizes soil acidity and increases fertility for improved plant growth. By altering soil pH, liming also influences net soil GHG emissions by microbial and non-microbial soil activity.

-   **other sources (including urea application, and non-energy products from fuels and solvent use)**. Urea-based fertilizers are applied to soils to improve the availability of nitrogen for plants. However increased nitrogen also impacts soil microbial activity, resulting in increased nitrous oxide emissions. Excess urea may leach into the environment or volatilize into ammonia as pollution. Non-energy products from fuels and solvent use include lubricants, paraffin waxes, bitumen/asphalt, and solvents. Carbon dioxide is the major greenhouse gas released from the use of these products.

The emission sources in this publication are groupings of categories in the Scottish Greenhouse Gas statistics. More information about how the emission sources have been grouped is available in the Methodology [LINK].

Some of these emission sources cut across several agriculture subsectors. For example, both dairy and arable farmers will produce emissions from the process of fuel combustion.

As reported in the Scottish Greenhouse Gas Statistics, enteric fermentation was the largest source of greenhouse gas emissions from agricultural activity in Scotland in 2022. Emissions from enteric fermentation were 4.0 MtCO~2~e, or just over half (52%) of total emissions from Scottish agriculture.

Looking at emission sources by subsector, just over half (52%) of emissions from enteric fermentation were produced by the suckler beef. Other major sources of emissions from agriculture include manure management which is mostly associated with livestock, and fuel combustion which is mostly associated with arable farming.

#### Figure 3: Enteric fermentation is the largest source of agriculture emissions

Agriculture greenhouse gas emissions in Scotland, by subsector and emission source, 2022

```{r fig3, echo = FALSE}
fig3

```

Whilst the largest source of agriculture emissions in 2022, enteric fermentation emissions have reduced by the largest amount. From 1990 and 2022, emissions fell by 0.8 MtCO~2~e (17%). The primary producers of emissions from enteric fermentation are cattle and sheep. Over this period, the number of cattle and sheep in Scotland reduced by 20% and 33%, respectively.

Emissions from liming and fuel combustion have both increased from 1990 to 2022 by a combined 0.4 MtCO~2~e. Arable farming accounts for most emissions from liming and fuel combustion.

# Farm GHG emissions

## Absolute gross emissions per hectare {.tabset}



### All farms

**Average farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

 
```{r fig5a, echo = FALSE, fig.height= 6}
fig5a

```


### Cereal

**Average cereal farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig5b, echo = FALSE, fig.height= 6}
fig5b

```


### General cropping

**Average general cropping farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig5c, echo = FALSE, fig.height= 6}
fig5c

```


### Dairy

**Average dairy farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig5d, echo = FALSE, fig.height= 6}
fig5d

```


### Specialist sheep (LFA)

**Average specialist sheep LFA (less favoured area) farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig5e, echo = FALSE, fig.height= 6}
fig5e

```


### Specialist cattle (LFA)

**Average specialist cattle LFA farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig5f, echo = FALSE, fig.height= 6}
fig5f

```

### Cattle and sheep (LFA)

**Average cattle and sheep LFA farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig5g, echo = FALSE, fig.height= 6}
fig5g

```


### Lowland cattle and sheep

**Average lowland cattle and sheep farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig5h, echo = FALSE, fig.height= 6}
fig5h

```

### Mixed

**Average mixed farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig5i, echo = FALSE, fig.height= 6}
fig5i

```
## {-}



Absolute emissions for the average Scottish farm were estimated at 4.4 tCO~2~e/ha in 2022-23. 

Emissions per hectare is a useful measure for arable farms where land area is more strongly related to output. In 2022-23, emissions for the average cereal farm were estimated at 2.8 tCO~2~e/ha. For general cropping, this was 2.2 tCO~2~e/ha. 

Emissions for the average cereal farm were more stable over the previous three years than those for general cropping farms. General cropping farms are likely to have more livestock and more intensive crops, e.g. potatoes, and have a different relationship between land area, output, and emissions.  

Dairy farms tend to have higher emissions per hectare than other livestock farms. This is due to higher stocking densities. In 2022-23, average dairy emissions were 17.8 tCO~2~e/ha. For the average lowland cattle and sheep farm (the next highest emitter), this was 6.9 tCO~2~e/ha.

Mixed farms represent a mix of crop and livestock activities and exhibit average emissions in between arable and livestock farm types. In 2022-23, average mixed farm emissions were 5.2 tCO~2~e/ha. 


## Emission intensity {.tabset}

### All farms 

**Average farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>
 
```{r fig6a, echo = FALSE, fig.height= 6}
fig6a

```


### Cereal

**Average cereal farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>
 
```{r fig6b, echo = FALSE, fig.height= 6}
fig6b

```


### General cropping

**Average general cropping farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>
 
```{r fig6c, echo = FALSE, fig.height= 6}
fig6c

```


### Dairy

**Average dairy farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>
 
```{r fig6d, echo = FALSE, fig.height= 6}
fig6d

```


### Specialist sheep (LFA)

**Average specialist sheep LFA (less favoured area) farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>
 
```{r fig6e, echo = FALSE, fig.height= 6}
fig6e

```


### Specialist cattle (LFA)

**Average specialist cattle LFA farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>
 
```{r fig6f, echo = FALSE, fig.height= 6}
fig6f

```

### Cattle and sheep (LFA)

**Average cattle and sheep LFA farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>
 
```{r fig6g, echo = FALSE, fig.height= 6}
fig6g

```


### Lowland cattle and sheep

**Average lowland cattle and sheep farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>
 
```{r fig6h, echo = FALSE, fig.height= 6}
fig6h

```

### Mixed

**Average mixed farm emissions greenhouse gas (GHG) emission intensity **

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>
 
```{r fig6i, echo = FALSE, fig.height= 6}
fig6i

```
## {-}




Emission intensity for the average farm was 7.0 kg CO~2~e/kg output in 2022-23. 

Farms vary widely in their emission intensity results, suggesting that there may be scope for some farms to decrease emission intensity. 

Livestock farms have higher emission intensities than arable or mixed farms. In 2022-23, sheep farms in less favoured areas (LFA) had the highest average emission intensity at 27.7 kg CO~2~e/kg output. 

Dairy farms, where the primary output is milk, have lower emission intensities compared with other livestock farms. The average dairy farm had an emission intensity of 1.5 kg CO~2~e/kg output in 2022-23. 

Cereals and general cropping farms had an average emission intensity of 0.4 and 0.3 kg CO~2~e/kg output respectively, in 2022-23. In the previous three years, both types remained similarly stable. 

Mixed farms showed an average emission intensity higher than arable farms, at 1.4 kg CO~2~e/kg. 



# Farm nitrogen use


## Nitrogen balance {.tabset}

### All farms

**Average farm nitrogen balance decreases**
 
 Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>
 
```{r fig7a, echo = FALSE, fig.height= 6}
fig7a

```


### Cereal

**Average cereal farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig7b, echo = FALSE, fig.height= 6}
fig7b

```


### General cropping

**Average general cropping farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig7c, echo = FALSE, fig.height= 6}
fig7c

```


### Dairy

**Average dairy farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig7d, echo = FALSE, fig.height= 6}
fig7d

```


### Specialist sheep (LFA)

**Average specialist sheep LFA (less favoured area) farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig7e, echo = FALSE, fig.height= 6}
fig7e

```


### Specialist cattle (LFA)

**Average specialist cattle LFA farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig7f, echo = FALSE, fig.height= 6}
fig7f

```

### Cattle and sheep (LFA)

**Average cattle and sheep LFA farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig7g, echo = FALSE, fig.height= 6}
fig7g

```


### Lowland cattle and sheep

**Average lowland cattle and sheep farm nitrogen balance increases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig7h, echo = FALSE, fig.height= 6}
fig7h

```

### Mixed farm

**Average mixed farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig7i, echo = FALSE, fig.height= 6}
fig7i

```
## {-}


Nitrogen balance on the average Scottish farm was 38.2 kg N surplus/ha in 2022-23, a decrease of 28% from 2021-22. A large nitrogen balance can indicate a higher risk of nitrogen being lost to the environment.  

On average, nitrogen output increased as farm outputs increased in 2022-23. Nitrogen inputs decreased, mostly because of decreased fertiliser input. 

Increases in output compared to the previous year drove the decrease in nitrogen balance.

Nitrogen from fertiliser makes the largest contribution to nitrogen input for general cropping and cereal farm types.

Estimates of the average nitrogen balance vary by farm type. 
Estimates of nitrogen balance on the average general cropping farm decreased dramatically (129%) in 2022-23 to -5.0 kg N surplus/ha.   A negative nitrogen balance is associated with a nitrogen use efficiency (NUE) that is above 100%. Both occur when nitrogen outputs are larger than nitrogen inputs. This may indicate a true loss of nitrogen from the soil and/or may be due to limitations of the methods used to capture all nitrogen inputs. 

The average cereal farm also showed a large decrease (66%) in nitrogen balance to 9.4 kg N surplus/ha.  Large decreases in fertiliser input and increases in output  in 2022-23 compared with the previous year drove the decreases in nitrogen balance.

Dairy farms have the highest average nitrogen balance of the farm types, estimated at 255.6 kg N surplus/ha in 2022-23. Nitrogen balance for the average dairy farm decreased 7% from the previous year. 

Lowland cattle and sheep were the only farm type where the average nitrogen balance increased in 2022-23, by 5% to 59.8 kg N surplus/ha.

Across the remaining livestock types, the average nitrogen balance decreased in 2022-23, most notably in LFA cattle and sheep, which decreased by 39% to 38.7 kg N surplus/ha. In LFA sheep and LFA cattle and sheep farms, decreases in nitrogen inputs drove this nitrogen balance decrease. In LFA cattle, increases in nitrogen outputs drove the decrease. 

Mixed farms have a nitrogen balance comparable to livestock farms, with an estimated average of 41.4 kg N surplus/ha in 2022-23, a decrease of 29% from previous years.  


## Nitrogen use efficiency {.tabset}

### All farms

**Average farm nitrogen use efficiency increases**
 
Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig8a, echo = FALSE, fig.height= 6}
fig8a

```


### Cereal

**Average cereal farm nitrogen use efficiency increases**
 
Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig8b, echo = FALSE, fig.height= 6}
fig8b

```


### General cropping

**Average general cropping farm nitrogen use efficiency increases**
 
Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig8c, echo = FALSE, fig.height= 6}
fig8c

```


### Dairy

**Average dairy farm nitrogen use efficiency remains stable**
 
Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig8d, echo = FALSE, fig.height= 6}
fig8d

```



### Specialist sheep (LFA)

**Average specialist sheep LFA (less favoured area) farm nitrogen use efficiency increases**
 
Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig8e, echo = FALSE, fig.height= 6}
fig8e

```


### Specialist cattle (LFA)

**Average specialist cattle LFA farm nitrogen use efficiency increases**
 
Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig8f, echo = FALSE, fig.height= 6}
fig8f

```

### Cattle and sheep (LFA)

**Average cattle and sheep LFA farm nitrogen use efficiency increases**
 
Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig8g, echo = FALSE, fig.height= 6}
fig8g

```


### Lowland cattle and sheep

**Average lowland cattle and sheep farm nitrogen use efficiency increases**
 
Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig8h, echo = FALSE, fig.height= 6}
fig8h

```

### Mixed

**Average mixed farm nitrogen use efficiency decreases**
 
Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. 
<div class="alert alert-info">
Double click the chart to zoom in. Double click again to reset.  
</div>

```{r fig8i, echo = FALSE, fig.height= 6}
fig8i

```
## {-}





# Glossary

# Data and Methodology

# Tell us what you think

We are always interested to hear from our users about how our statistics are used, and how they can be improved.

## Feedback survey

We'd appreciate it if you would complete our short feedback survey on this publication.

## Enquiries

For enquiries about this publication please contact:

[Name(s) of lead statistician(s)]

[Department name]

E-mail: [e-mail address]

Telephone: [Phone number]

For general enquiries about Scottish Government statistics please contact:

Office of the Chief Statistician

e-mail: statistics.enquiries\@gov.scot

## Join our mailing list

If you would like to receive notification about statistical publications, or find out about consultations on our statistics please join the [ScotStat mailing list](http://www.gov.scot/scotstat).

## Future publications

Details of future publications can be found on our [forthcoming publications](https://www.gov.scot/publications/official-statistics-forthcoming-publications/) page.

