---
title: "Scottish agriculture greenhouse gas emissions and nitrogen use: 2023-24"
date: "10 June 2025"
navigation:
    toc:
      sticky: true
output:
  html_document:
    template: _template.html
    toc: true
    toc_depth: 1
    toc_float: false
    self_contained: false
sgtemplates:
  footer:
    link:
      - href: "accessibility.html"
        text: "Accessibility"
      - href: "contact.html"
        text: "Contact"
  header:
    site_branding: "Scottish agriculture greenhouse gas emissions and nitrogen use: 2023-24"

seo:
  seo:
    description: "SEO (search engine optimisation) description"
    keywords: "agriculture 2023, agriculture emissions, scotland ghg, beef emissions"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
options(knitr.kable.NA = '')
source("scripts/highcharter_fig1-3.R")
source("scripts/gross_emissions_figs.R")
source("scripts/nitrogen_figs.R")
source("scripts/html tables.R")

```

::: {.alert .alert-info}
Part of: <a href ="https://www.gov.scot/collections/scottish-agriculture-greenhouse-gas-emissions-and-nitrogen-use/" target ="_blank">Scottish agriculture greenhouse gas emissions and nitrogen use</a>

Find more <a href ="https://www.gov.scot/collections/agriculture-fisheries-and-rural-statistics/" target ="_blank">agriculture, fisheries and rural statistics</a>

Visit our <a href ="https://scotland.shinyapps.io/sg-agric-stats-hub/" target ="_blank">Agricultural Statistics Hub</a> for more data and interactive charts
:::


<b>An official statistics in development publication for Scotland </b>

Publication date: 10 June 2025

<a href = download/scottish_agriculture_ghg_n_use_data_table.xlsx> <button class="btn noprint"> Download data in .xlsx format </button></a>


This report is based on estimates of total greenhouse gas (GHG) emissions for Scottish agriculture, as reported in the <a href ="https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2023/" target ="_blank">Scottish Greenhouse Gas Statistics 2023</a> and average emission and nitrogen use estimates from the Farm Business Survey (FBS). The report includes:

-   Results for total Scottish agriculture <abbr title="greenhouse gas">GHG</abbr> emissions split by **agricultural subsector** (e.g. suckler beef, arable, dairy etc). These subsectors align more closely with agricultural activity and provide a different breakdown to the categories used in the Scottish Greenhouse Gas Statistics.

-   **Average whole farm emissions of different farm types** (e.g. lowland cattle and sheep, general cropping, LFA (less favoured area) sheep etc). Farms are assigned to a farm type based on their primary output. The results also show the variation in emissions within each farm type. This is based on FBS data and is weighted to cover commercial sized farms in Scotland.

-   **Average emission intensity of specific agricultural outputs of different farm types.**  For example, estimates for the average emission intensity of a kg deadweight (dwt) of beef in an LFA cattle farm, a lowland cattle and sheep farm and an LFA cattle and sheep farm. Agricultural outputs are the products of farm enterprises – agricultural activities within a farm business. Estimates are not weighted because they are based on a sub-sample of data.  

-   **Average farm level nitrogen use across different farm types.** Estimates are weighted using the same methodology described above.

-   Analysis of whole farm emissions, emission intensity of agricultural outputs and nitrogen use are based on farms in the FBS. This is a survey of around 400 commercial farms (whose standard output exceeds around £20,000). The sample excludes farm types not in receipt of support payments (such as pigs, poultry and horticulture). Some results are weighted to better reflect the wider population. Some results are presented unweighted because they are based on sub-samples of data. Where the sample sizes are small, data can be more highly variable and have a wider margin of uncertainty. 

More information is available in the <a href = methodology.html> Methodology</a>.

Information about the different data sources and coverage of the estimates in the report is displayed in the table below. 

<button class="btn btn-primary" data-toggle="collapse" data-target="#Table1">
  
Show/Hide information table

</button>
  
::: {#Table1 .collapse}
```{r table_1_html, echo = FALSE, results = 'markup'}
knitr::kable(table_1_html, format = "html", escape = FALSE) %>% kableExtra::kable_styling(font_size = 16,  html_font = 'Arial', bootstrap_options= 'responsive')
```
:::

<br></br>


# Key Points

<div style="margin-top: 30px;"></div>

## Agricultural sector and subsectors

-   In 2023, <abbr title="greenhouse gas">GHG</abbr> emissions from Scottish agriculture were 7.5 million tonnes of carbon dioxide equivalent (MtCO~2~e). 

-   Scottish agriculture emissions were stable in 2023, falling by 0.1% from 2022 to their lowest level since current records began in 1990. Small increases in emissions from fuel combustion, liming and urea application were narrowly offset by falls in emissions from enteric fermentation, agricultural soils and manure. <a href ="https://assets.publishing.service.gov.uk/media/6697b450fc8e12ac3edafe7b/BSFP_2023_digital_bookmarked.pdf" target ="_blank">Fertiliser usage rates increased slightly in 2023, as market prices lowered from the peak of 2022</a>. Fuel prices also fell from the high of 2022.

-   Total agriculture emissions have been in a long-term decline. Since 1990 emissions have fallen by 13% from 8.6 MtCO~2~e to 7.5 MtCO~2~e. Over the same period, total GHG emissions in Scotland fell by 51% from 81.2 MtCO~2~e to 39.6 MtCO~2~e.

-   The long-term fall in emissions from agriculture was largely driven by falling livestock numbers. From 1990 to 2023 cattle and sheep numbers have decreased by 20% and 33%, respectively.

-   Arable and other subsectors (including pigs, poultry and non-agriculture) were the only subsectors with increased emissions in 2023. Arable farming emissions rose by 5% to 1.5 MtCO~2~e, driven by rising emissions from fuel combustion and liming. Other subsector emissions rose by 1% to 0.2 MtCO~2~e.

-   The suckler beef subsector is the largest contributor to total agriculture emissions. In 2023 suckler beef emissions contributed 41% of total agriculture emissions. Suckler beef emissions fell by 0.7 MtCO~2~e (18%) between 1990 and 2023. 

<div style="margin-top: 30px;"></div>

## Average farm emissions

-   Emission intensity results are available for the first time for farms in the Farm Business Survey (FBS). Average emission intensity measures the carbon footprint of a specific product from an activity or enterprise on a farm. For example, the average amount of GHG emitted in the process of producing a tonne of wheat, or a kilogram of beef.

-   In 2023-24 average emission intensity for cereals production increased on both cereal (by 14% to 258 kgCO~e~/tonne crop) and general cropping (by 7% to 241 kgCO~2~e/tonne crop) farms in the survey compared to the previous year. The rise was mostly driven by increased emissions from fertiliser and manure. Fertiliser usage rates rose in 2023-24 as prices fell from their peak in 2022-23.

-   Across the livestock farm types included in the Farm Business Survey, average beef emission intensity ranged from 30.9 to 32.8 kgCO~2~e/kg dwt in 2023-24.

-   Average sheep emission intensity was higher on Less Favoured Area (LFA) sheep farms (35.5 kgCO~2~e/kg dwt) than on lowland sheep farms (25.2 kgCO~2~e/kg dwt). Average sheep output was similar, despite considerably lower average herd sizes on lowland farms in 2023-24. This may indicate a higher productivity on lowland farms. 

-   The average emission intensity for milk production was 1.3 kgCO~2~e/kg fat and protein corrected (FPC) milk in 2023-24. This was an increase of 2% from the previous year, as average milk yields fell. Lower productivity is associated with higher emission intensities.

<div style="margin-top: 30px;"></div>

## Average farm nitrogen use and efficiency

-   Nitrogen balance, the difference between nitrogen inputs and outputs, on the average Scottish farm was 45.9 kg N surplus/ha in 2023-24. This was an increase of 20% from the previous year and may indicate that more nitrogen from agriculture was at risk of being lost to the environment. Despite the increase from 2022-23, nitrogen balance in 2023-24 was at the second lowest value since the timeseries began in 2019-20. 

-   Nitrogen balance rose for most farm types on average in 2023-24. In general, the rise in nitrogen balance was driven by increased fertiliser and high energy feed inputs. Fertiliser rates rose slightly in 2023-24 as usage rates recovered from the previous year. This follows the trend seen in the GHG emissions results. Falling cereal outputs, where lower yields can lead to nitrogen accumulation in the soil, also drove increases in nitrogen balance.

-   Average nitrogen balance fell for both LFA cattle and lowland cattle and sheep farms in 2023-24. Decreases were driven by falling fertiliser and low energy concentrate feed inputs or a rise in inputs with high nitrogen availability, such as bean and draff. 

-   Nitrogen use efficiency (NUE), the ratio of nitrogen outputs to inputs, was estimated at 29% for the average farm in 2023-24. This is a decrease of two percentage points compared with 2022-23 (31%). A higher NUE typically indicates a more efficient use of nitrogen. 

<br></br>

# Introduction

This report provides GHG emissions for the Scottish agriculture sector, agricultural subsectors and for average Scottish farms by farm type. <abbr title="agricultural activities within a farm business">Enterprise</abbr> level estimates provide the average carbon footprint of products produced by farms in the Farm Business survey. Nitrogen use estimates are also included for average Scottish farms. Several data sources have been combined to provide a broad view of emissions and nitrogen use from the agriculture sector. More information is available in the [about the statistics section](#about-statistics).

Farm-level gross emission results are per hectare. This helps with comparing businesses of different sizes. Emission intensity results are for the primary product of each farm enterprise (activity). For beef and sheep enterprises, emission intensities are calculated per kg of carcass deadweight (kg dwt). Milk emission intensities for dairy enterprises are calculated per kg of fat and protein corrected milk (kg FPC milk).  Cereals emission intensities are calculated for total barley, wheat, oats and minor cereals enterprises (kg tonnes). 

Methodology differences mean average farm estimates are not directly comparable with national estimates. Methods differ in their coverage of the industry and emissions. 

Estimates cover similar but different time periods. Emissions for the agriculture sector and its subsectors are based on a calendar year, whilst farm-level results are based on an accounting year. 

It is important to note that agriculture and food production rely on natural processes. This means that agricultural activity will always cause a degree of GHG emissions. Emissions will also vary widely between different agricultural produce. 

<div style="margin-top: 30px;"></div>

## An official statistics in development publication for Scotland

These statistics are official statistics in development. Official statistics in development may be new or existing statistics, and will be tested with users, in line with the standards of trustworthiness, quality, and value in the <a href ="https://code.statisticsauthority.gov.uk/the-code/" target ="_blank">Code of Practice for Statistics</a>.

Scottish Government statistics are regulated by the Office for Statistics Regulation (OSR). OSR sets the standards of trustworthiness, quality and value in the <a href ="https://code.statisticsauthority.gov.uk/the-code/" target ="_blank">Code of Practice for Statistics</a> that all producers of official statistics should adhere to.

<br></br>

# Agriculture sector GHG emissions {#agsector}

<div style="margin-top: 40px;"></div>

#### Figure 1: Greenhouse gas (GHG) emissions from agriculture at lowest levels since 1990

Total and agriculture GHG emissions in Scotland, 1990 to 2023

::: {.alert .alert-info}
Click and drag to zoom in on part of the chart. Click "Reset zoom" to reset.
:::

```{r fig1, echo = FALSE}
fig1

```

<font size="3"><strong>Source:</strong> <a href ="https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2023/" target ="_blank">Scottish Greenhouse Gas Statistics 2023</a>

<strong> Download data for this chart </strong>

<a href = download/Agriculture_emissions.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/Agriculture_emissions.csv> <button class="btn pdf"> .csv </button> </a>

<b> Estimates for total Scottish and agriculture greenhouse gas emissions have been revised since last year’s release to incorporate methodological improvements and new data. </b> Improvements to data used to estimate Scotland’s share of UK mobile machinery emissions have been applied to the entire timeseries. As a result, total Scottish and agriculture emissions are slightly lower than previously reported.  

GHG emissions from Scottish agriculture were 7.5 million tonnes of carbon dioxide equivalent (MtCO~2~e) in 2023. This is the lowest value since the baseline period in 1990, a decrease of 1.1 MtCO~2~e over this period. 

Since 1990 (8.6 MtCO~2~e), total agriculture emissions decreased by 13%. Total Scotland emissions have fallen by 51% over the same period, from 81.2 MtCO~2~e in 1990 to 39.6 MtCO~2~e in 2023. In 2023, Agriculture was the third largest source of emissions in Scotland behind Domestic transport and Buildings and product uses, using Territorial Emissions Statistics Sector categories.

Total agriculture emissions in 2023 remained relatively stable, falling by 0.1% from the previous year (from 7.51 MtCO~2~e in 2022 to just under this in 2023). Decreases in emissions from enteric fermentation (-2%, 0.08 MtCO~2~e), agricultural soils (-2%, 0.03 MtCO~2~e) and manure (-2%, 0.02 MtCO~2~e) narrowly offset increases in emissions from fuel combustion (10%, 0.08 MtCO~2~e), liming (7%, 0.02 MtCO~2~e) and urea application (113%, 0.02 MtCO~2~e). In 2023, <a href ="https://assets.publishing.service.gov.uk/media/6697b450fc8e12ac3edafe7b/BSFP_2023_digital_bookmarked.pdf" target ="_blank">fertiliser usage rates increased slightly, as market prices lowered from the peak of 2022</a>. <a href ="https://ahdb.org.uk/fuel-prices" target ="_blank">Fuel prices also fell from the high of 2022,</a> as markets were recovering following Russia’s invasion of Ukraine. 

More results for total Scotland GHG emissions, including for agriculture, are available in the <a href ="https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2023/" target ="_blank">Scottish Greenhouse Gas (GHG) Statistics 2023</a>.

<br></br>

# Agriculture subsector GHG emissions

<div style="margin-top: 30px;"></div>

#### Figure 2: Suckler beef has the largest share of Scottish agriculture greenhouse gas (GHG) emissions

Agriculture GHG emissions in Scotland, by subsector, 2023

::: {.alert .alert-info}
Click and drag to zoom in on part of the chart. Click "Reset zoom" to reset.
:::

```{r fig2, echo = FALSE}
fig2

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<strong> Download data for this chart </strong>

::: {style="margin-bottom: 20px;"}
<a href = download/Subsector_emissions.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/Subsector_emissions.csv> <button class="btn pdf"> .csv </button> </a>
:::

The suckler beef subsector had the largest share of Scottish agriculture emissions from 1990 to 2023. In 2023 suckler beef accounted for 41% (3.1 MtCO~2~e) of total agricultural emissions. The second largest contribution is from arable (19% or 1.5 MtCO~2~e in 2023), closely followed by dairy (18% or 1.3 MtCO~2~e in 2023) and sheep (16% or 1.2 MtCO~2~e in 2023). Dairy beef and other subsectors (including pigs, poultry, and non-agriculture) accounted for a small proportion (3% each) of total agriculture emissions in 2023.  

In 2023, suckler beef, dairy, sheep and dairy beef emissions fell by 1% each, compared with the previous year. Arable and other (including pigs, poultry, and non-agriculture) were the only subsectors in 2023 with increased emissions from the previous year, increasing by 5% and 1% respectively. A rise in machinery and fertiliser use, both of which make large contributions to arable farming, drove this increase.  

Reductions in emissions from suckler beef have contributed to the largest absolute decrease in total agriculture emissions since the baseline period. From 1990 to 2023 suckler beef emissions fell by 0.7 MtCO~2~e (18%). This was driven by declining cattle numbers over the same period, from 2.1 million in 1990 to 1.7 million in 2023, a decrease of 20%. The suckler beef subsector’s contribution to total agriculture emissions fell from 44% in 1990 to 41% in 2023.

Emissions from arable farming, which has the next largest share of total emissions, increased by 15% from 1990 to 2023. This the only subsector with increased emissions compared with 1990, driven by increased emissions from machinery and fertiliser use.  

Dairy and sheep emissions, moderate contributors to total agriculture emissions, fell by 0.3 MtCO~2~e (17% and 19% respectively) from 1990 to 2023.     

<div style="margin-top: 40px;"></div>

## Sources of agriculture GHG emissions

#### Figure 3: Enteric fermentation is the largest source of agriculture emissions

Agriculture greenhouse gas (GHG) emissions in Scotland, by subsector and emission source, 2023

::: {.alert .alert-info}
Click and drag to zoom in on part of the chart. Click "Reset zoom" to reset. Click on individual emission sources on the right to select or deselect.
:::

```{r fig3, echo = FALSE}
fig3

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<strong> Download data for this chart </strong>

::: {style="margin-bottom: 20px;"}
<a href = download/Subsector_source_emissions.xlsx> <button class="btn pdf"> .xlsx </button> </a><a href = download/Subsector_source_emissions.csv> <button class="btn pdf"> .csv </button></a>
:::

As reported in the <a href ="https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2023/" target ="_blank">Scottish Greenhouse Gas (GHG) Statistics 2023</a>, enteric fermentation (animal digestion) was the largest source of GHG emissions from agriculture in 2023. Emissions from enteric fermentation were 4.0 MtCO~2~e, just over half (53%) of total emissions from Scottish agriculture. 

Looking at emission sources by subsector, just over half (52%) of emissions from enteric fermentation were produced by suckler beef. 

Emissions from enteric fermentation have had the largest overall reduction compared with all other agricultural emission sources. From 1990 and 2023, emissions from enteric fermentation fell by 0.9 MtCO~2~e (19%). The primary producers of emissions from enteric fermentation are cattle and sheep. Over this period, the number of cattle and sheep in Scotland reduced by 20% and 33%, respectively.

Emissions from liming and fuel combustion have both increased between 1990 and 2023 by a combined 0.4 MtCO~2~e (51%). Arable farming accounts for most emissions from liming and fuel combustion.

Another major source of emissions from agriculture is manure management, which is mostly associated with livestock. Manure management emissions fell by 25% or 0.5 MtCO~2~e from 1990 to 2023. 


<br></br>

# Whole farm GHG emissions

Absolute gross emissions per hectare estimates are weighted to the June Agricultural Census population. <abbr title="agricultural activities within a farm business">Enterprise</abbr> level emission intensity estimates are not weighted because these are based on a sub-sample of data. These replace previously published whole-farm emission intensities, which were weighted and measured the combined carbon footprint of all products on a farm.

::: {.alert .alert-info}
We are very interested to hear your opinions on our application of weighting methodology and the presentation of these data. Please let us know your thoughts by filling in this quick <a href ="https://forms.office.com/e/2SCLV7aTJi" target ="_blank">survey</a>. 
:::

<div style="margin-top: 40px;"></div>

## Absolute gross emissions per hectare {.tabset}

### All farms

**Figure 4a: Average farm absolute gross greenhouse gas (GHG) emissions remain stable**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval. 

<font size="3">Due to methodology improvements, results for 2021-22 onwards are not directly comparable to previous years. Previous years will not be revised as not all data are available.</font>

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5a, echo = FALSE, fig.height=5.5, fig.width=8}
fig5a
```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Absolute_AllTypes.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/Absolute_AllTypes.csv> <button class="btn pdf"> .csv </button> </a>

### Cereal

**Figure 4b: Average cereal farm absolute gross greenhouse gas (GHG) emissions slightly decrease**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval. 

<font size="3">Due to methodology improvements, results for 2021-22 onwards are not directly comparable to previous years. Previous years will not be revised as not all data are available.</font>

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5b, echo = FALSE, fig.height=5.5, fig.width=8}
fig5b
```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24.</font>

<b> Download data for this chart </b>

<a href = download/Absolute_Cereal.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/Absolute_Cereal.csv> <button class="btn pdf"> .csv </button> </a>

### General cropping

**Figure 4c: Average general cropping farm absolute gross greenhouse gas (GHG) emissions increase**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

<font size="3">Due to methodology improvements, results for 2021-22 onwards are not directly comparable to previous years. Previous years will not be revised as not all data are available.</font>

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5c, echo = FALSE, fig.height=5.5, fig.width=8}
fig5c
```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24.</font>

<b> Download data for this chart </b>

<a href = download/Absolute_GeneralCropping.xlsx > <button class="btn pdf"> .xlsx </button> </a> <a href = download/Absolute_GeneralCropping.csv > <button class="btn pdf"> .csv </button> </a>

### Dairy

**Figure 4d: Average dairy farm absolute gross greenhouse gas (GHG) emissions decrease**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval. 

<font size="3">Due to methodology improvements, results for 2021-22 onwards are not directly comparable to previous years. Previous years will not be revised as not all data are available.</font>

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5d, echo = FALSE, fig.height=5.5, fig.width=8}
fig5d
```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Absolute_Dairy.xlsx > <button class="btn pdf"> .xlsx </button> </a> <a href = download/Absolute_Dairy.csv > <button class="btn pdf"> .csv </button> </a>

### LFA sheep

**Figure 4e: Average LFA (less favoured area) sheep farm absolute gross greenhouse gas (GHG) emissions increase**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval. 

<font size="3">Due to methodology improvements, results for 2021-22 onwards are not directly comparable to previous years. Previous years will not be revised as not all data are available.</font>

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5e, echo = FALSE, fig.height=5.5, fig.width=8}
fig5e
```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Absolute_LFASheep.xlsx > <button class="btn pdf"> .xlsx </button> </a> <a href = download/Absolute_LFASheep.csv > <button class="btn pdf"> .csv </button> </a>

### LFA cattle

**Figure 4f: Average LFA (less favoured area) cattle farm absolute gross greenhouse gas (GHG) emissions decrease**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

<font size="3">Due to methodology improvements, results for 2021-22 onwards are not directly comparable to previous years. Previous years will not be revised as not all data are available.</font>

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5f, echo = FALSE, fig.height=5.5, fig.width=8}
fig5f
```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Absolute_LFACattle.xlsx > <button class="btn pdf"> .xlsx </button> </a> <a href = download/Absolute_LFACattle.csv > <button class="btn pdf"> .csv </button> </a>

### LFA cattle and sheep

**Figure 4g: Average LFA (less favoured area) cattle and sheep farm absolute gross greenhouse gas (GHG) emissions increase**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval. 

<font size="3">Due to methodology improvements, results for 2021-22 onwards are not directly comparable to previous years. Previous years will not be revised as not all data are available.</font>

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5g, echo = FALSE, fig.height=5.5, fig.width=8}
fig5g
```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Absolute_LFACattleandSheep.xlsx > <button class="btn pdf"> .xlsx </button> </a> <a href = download/Absolute_LFACattleandSheep.csv > <button class="btn pdf"> .csv </button> </a>

### Lowland cattle and sheep

**Figure 4h: Average lowland cattle and sheep farm absolute gross greenhouse gas (GHG) emissions decrease**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval. 

<font size="3">Due to methodology improvements, results for 2021-22 onwards are not directly comparable to previous years. Previous years will not be revised as not all data are available.</font>

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5h, echo = FALSE, fig.height=5.5, fig.width=8}
fig5h
```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Absolute_LowlandCattleandSheep.xlsx > <button class="btn pdf"> .xlsx </button> </a> <a href = download/Absolute_LowlandCattleandSheep.csv > <button class="btn pdf"> .csv </button> </a>

### Mixed

**Figure 4i: Average mixed farm absolute gross greenhouse gas (GHG) emissions decrease**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval. 

<font size="3">Due to methodology improvements, results for 2021-22 onwards are not directly comparable to previous years. Previous years will not be revised as not all data are available.</font>

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5i, echo = FALSE, fig.height= 5.5, fig.width= 8}
fig5i
```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Absolute_Mixed.xlsx > <button class="btn pdf"> .xlsx </button> </a> <a href = download/Absolute_Mixed.csv > <button class="btn pdf"> .csv </button> </a>

##  {.unnumbered}

Absolute emissions for the average Scottish farm were estimated at 4.2 tCO~2~e/ha in 2023-24. 

Emissions per hectare is a useful measure for arable farms where land area is more strongly related to output. In 2023-24, emissions for the average general cropping farm rose by 32% to 2.1 tCO~2~e/ha. The rise was driven by an increase in emissions from fertiliser and manure use. Average cereal farm emissons was estimated at 2.0 tCO~2~e/ha, a slight decrease of 3% compared with emissions in 2022-23. Whilst fertiliser and manure use increased on average cereal farms in 2023-24, the fall in absolute emissions was driven by a small increase in land area on the average cereal farm. 

Emissions for the average cereal farm are more stable than those for general cropping farms. General cropping farms tend to have more livestock and more intensive crops, e.g. potatoes. They can have a different relationship between land area, output, and emissions.  

Dairy farms tend to have higher emissions per hectare than other livestock farms. This is due to higher stocking densities. In 2023-24, average dairy emissions were 16.2 tCO~2~e/ha, a 1% decrease from emissions in 2023-24. Despite higher average herd sizes, the fall in dairy emissions in 2023-24 was mainly driven by a 30% increase in land area on the average dairy farm.   

Average emissions from Less Favoured Area (LFA) cattle farms were 5.5 tCO~2~e/ha in 2023-24, a fall of 6% compared with the previous year. The fall in total emissions was largely driven by a decrease in average emissions from methane. 

In 2023-24, average lowland cattle and sheep farm emissions also decreased, falling by 11% from the previous year to 5.4 tCO~2~e/ha. 

Other livestock farm types saw an increase in average emissions in 2023-24. Compared with the previous year, average emissions for LFA sheep farms rose by 6% to 3.5 tCO~2~e/ha. Increases in nitrous oxide emissions associated with grazing, manure spreading, manure management and crop residues largely drove the rise in total average emissions from LFA sheep farms in 2023-24. 

Average LFA cattle and sheep farm emissions rose by 12% to 4.3 tCO~2~e/ha in 2023-24, compared with the previous year. Increases in nitrous oxide emissions and a fall in land area for the average farm drove this increase.

Results for average LFA sheep and LFA cattle and sheep farms are partly impacted by sample turnover. In 2023-24, LFA sheep and LFA cattle farms entering the sample had higher gross emissions and smaller land areas on average than farms of the same farm type that left the sample. 

Mixed farms represent a mix of crop and livestock activities and exhibit average emissions in between arable and livestock farm types. In 2023-24, average mixed farm emissions were 3.7 tCO~2~e/ha, a decrease of 16% from the previous year. 

<br></br>

# Enterprise GHG emission intensity

Emission intensities (kgCO~2~e/kg product) measure the carbon footprint of a farm enterprise (activity). A lower emission intensity can indicate a more efficient process used to produce a specific output.

Average enterprise level emission intensity estimates are not weighted to the 2023 June Agricultural Census population. This means that results do not represent the population, but indicate average emissions from specific farm activities by a sample of farms in Scotland. 

::: {.alert .alert-info}
We are very interested to hear your opinions on our application of weighting methodology and the presentation of these data. Please let us know your thoughts by filling in this quick <a href ="https://forms.office.com/e/2SCLV7aTJi" target ="_blank">survey</a>. 
:::

<div style="margin-top: 40px;"></div>

## Beef emission intensity

**Table 2: Average cattle livestock farm beef emission intensities (kgCO~2~e/kg dwt) in the Farm Business Survey 2023-24 sample**

<button class="btn btn-primary" data-toggle="collapse" data-target="#Table2">
  
Show/Hide data table

</button>
  
::: {#Table2 .expand}
```{r table_2_html, echo = FALSE, results = 'markup'}
knitr::kable(table_2_html) %>% kableExtra::kable_styling(font_size = 16,  html_font = 'Arial', full_width = FALSE, position = "left", bootstrap_options= 'responsive')
```
:::
      
<strong> Download data for this table </strong>
      
::: {style="margin-bottom: 20px;"}
<a href = download/Beef_enterprise.xlsx> <button class="btn pdf"> .xlsx </button> </a><a href = download/Beef_enterprise.csv> <button class="btn pdf"> .csv </button></a>
:::

In 2023-24 there was little variation between average beef emission intensity estimates for cattle livestock farm types in the FBS. LFA (less favoured area) cattle and sheep farms had the largest average beef emission intensity at 32.8 kgCO2e/kg dwt. Beef emission intensities for LFA cattle and lowland cattle and sheep were 31.8 kgCO~2~e/kg dwt and 30.9 kgCO~2~e/kg dwt, respectively. 

Livestock emission intensities can indicate how efficient the conversion of feedstuffs into liveweight gains is. Breeding selection, animal health and dietary management are factors that can influence the rate of enteric fermentation. Enteric fermentation is the major source of emissions from livestock enterprises. 

<div style="margin-top: 40px;"></div>

## Sheep emission intensity

**Table 3: Average sheep livestock farm sheep emission intensities (kgCO~2~e/kg dwt) in the Farm Business Survey 2023-24 sample**

<button class="btn btn-primary" data-toggle="collapse" data-target="#Table3">
  
Show/Hide data table

</button>
  
::: {#Table3 .expand}
```{r table_3_html, echo = FALSE, results = 'markup'}
knitr::kable(table_3_html) %>% kableExtra::kable_styling(font_size = 16,  html_font = 'Arial', full_width = FALSE, position = "left", bootstrap_options= 'responsive')
```
:::
      
<strong> Download data for this table </strong>
      
::: {style="margin-bottom: 20px;"}
<a href = download/Sheep_enterprise.xlsx> <button class="btn pdf"> .xlsx </button> </a><a href = download/Sheep_enterprise.csv> <button class="btn pdf"> .csv </button></a>
:::

In 2023-24 LFA sheep farms in the FBS sample had the highest average sheep emission intensity at 35.5 kgCO~2~e/kg dwt. Lowland cattle and sheep farms had the lowest average sheep emission intensity at 25.2 kgCO~2~e/kg dwt in 2023-24. 

Average sheep output in 2023-24 was similar for both farm types, despite considerably lower average herd sizes on lowland farms in 2023-24. This may indicate a higher productivity on lowland farms compared to LFA farms. 

<div style="margin-top: 40px;"></div>

## Milk emission intensity

**Table 4: Average dairy farm milk emission intensities (kgCO~2~e/kg fat and protein corrected (FPC) milk) in the Farm Business Survey sample 2022-23 and 2023-24**

<button class="btn btn-primary" data-toggle="collapse" data-target="#Table4">
  
Show/Hide data table

</button>
  
::: {#Table4 .expand}
```{r table_4_html, echo = FALSE, results = 'markup'}
knitr::kable(table_4_html, align = "lrrr") %>% kableExtra::kable_styling(font_size = 16,  html_font = 'Arial', full_width = FALSE, position = "left", bootstrap_options= 'responsive') 
```
:::
      
<strong> Download data for this table </strong>
      
::: {style="margin-bottom: 20px;"}
<a href = download/Milk_enterprise.xlsx> <button class="btn pdf"> .xlsx </button> </a><a href = download/Milk_enterprise.csv> <button class="btn pdf"> .csv </button></a>
:::

Average milk emission intensity on dairy farms in the FBS was 1.3 kgCO~2~e/kg FPC milk in 2023-24. This was an increase of 2% from the previous year. In 2023-24 average milk yields were lower. Lower productivity is associated with higher emission intensities. Milk emission intensities are also influenced by farm management practices such as feed efficiency, age at first calving and whether herds are grazing, housed or mixed.   

<div style="margin-top: 40px;"></div>

## Cereals emission intensity

**Table 5: Average arable farm cereal emission intensities (kgCO~2~e/tonne crop) in the Farm Business Survey sample 2022-23 and 2023-24**

<button class="btn btn-primary" data-toggle="collapse" data-target="#Table5">
  
Show/Hide data table

</button>
  
::: {#Table5 .expand}
```{r table_5_html, echo = FALSE, results = 'markup'}
knitr::kable(table_5_html, align = "lrrr") %>% kableExtra::kable_styling(font_size = 16,  html_font = 'Arial', full_width = FALSE, position = "left", bootstrap_options= 'responsive')
```
:::
      
<strong> Download data for this table </strong>
      
::: {style="margin-bottom: 20px;"}
<a href = download/Cereals_enterprise.xlsx> <button class="btn pdf"> .xlsx </button> </a><a href = download/Cereals_enterprise.csv> <button class="btn pdf"> .csv </button></a>
:::

In 2023-24 cereal farms had the highest average cereals emission intensity at 258 kgCO~2~e/ tonne crop. This was an increase of 14% from the previous year. Average cereals emission intensity for general cropping farms rose by 7% to 241 kgCO~2~e/tonne crop over the same period. The rise in average cereals emission intensities was mostly driven by increased emissions from fertiliser and manure.  Fertiliser usage rates rose slightly in 2023-24 as market prices fell from their peak in 2022-23.

<br></br>

# Farm nitrogen use {#n-balance}
<div style="margin-top: 40px;"></div>
## Nitrogen balance {.tabset}

### All farms

**Figure 5a: Average farm nitrogen balance increases**

Median farm nitrogen balance, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6a, echo = FALSE, fig.height=5.5, fig.width=8}
fig6a

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Nbalance_AllTypes.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/Nbalance_AllTypes.csv> <button class="btn pdf"> .csv </button> </a>

### Cereal

**Figure 5b: Average cereal farm nitrogen balance increases**

Median farm nitrogen balance, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6b, echo = FALSE, fig.height=5.5, fig.width=8}
fig6b

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Nbalance_Cereal.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/Nbalance_Cereal.csv> <button class="btn pdf"> .csv </button> </a>

### General cropping

**Figure 5c: Average general cropping farm nitrogen balance increases**

Median farm nitrogen balance, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6c, echo = FALSE, fig.height=5.5, fig.width=8}
fig6c

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Nbalance_GeneralCropping.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/Nbalance_GeneralCropping.csv> <button class="btn pdf"> .csv </button> </a>

### Dairy

**Figure 5d: Average dairy farm nitrogen balance increases**

Median farm nitrogen balance, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6d, echo = FALSE, fig.height=5.5, fig.width=8}
fig6d

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Nbalance_Dairy.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/Nbalance_Dairy.csv> <button class="btn pdf"> .csv </button> </a>

### LFA sheep

**Figure 5e: Average LFA (less favoured area) sheep farm nitrogen balance increases**

Median farm nitrogen balance, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6e, echo = FALSE, fig.height=5.5, fig.width=8}
fig6e

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Nbalance_LFASheep.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/Nbalance_LFASheep.csv> <button class="btn pdf"> .csv </button> </a>

### LFA cattle

**Figure 5f: Average LFA (less favoured area) cattle farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6f, echo = FALSE, fig.height=5.5, fig.width=8}
fig6f

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Nbalance_LFACattle.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/Nbalance_LFACattle.csv> <button class="btn pdf"> .csv </button> </a>

### LFA cattle and sheep

**Figure 5g: Average LFA cattle and sheep farm nitrogen balance increases**

Median farm nitrogen balance, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6g, echo = FALSE, fig.height=5.5, fig.width=8}
fig6g

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Nbalance_LFACattleandSheep.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/Nbalance_LFACattleandSheep.csv> <button class="btn pdf"> .csv </button> </a>

### Lowland cattle and sheep

**Figure 5h: Average lowland cattle and sheep farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6h, echo = FALSE, fig.height=5.5, fig.width=8}
fig6h

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Nbalance_LowlandCattleandSheep.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/Nbalance_LowlandCattleandSheep.csv> <button class="btn pdf"> .csv </button> </a>

### Mixed farm

**Figure 5i: Average mixed farm nitrogen balance increases**

Median farm nitrogen balance, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6i, echo = FALSE, fig.height=5.5, fig.width=8}
fig6i

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/Nbalance_Mixed.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/Nbalance_Mixed.csv> <button class="btn pdf"> .csv </button> </a>

##  {.unnumbered}

Nitrogen balance measures the difference between nitrogen input and nitrogen output, accounting for land area.

Nitrogen balance on the average Scottish farm was 45.9 kg N surplus/ha in 2023-24, an increase of 20% from 2022-23. Despite the increase from 2022-23, nitrogen balance in 2023-24 was at the second lowest value since the timeseries began in 2019-20. A large nitrogen balance can indicate a higher risk of nitrogen being lost to the environment. More information about nitrogen estimates is available here.  

For the average farm, the decrease in nitrogen balance was mostly driven by increases in nitrogen input. Nitrogen input increased as farm output decreased in 2023-24. A rise in fertiliser input mostly drove the increase in nitrogen inputs in 2023-24. Fertiliser usage rates rose slightly in 2023-24 as market prices fell from their peak in 2022-23.  

Estimates of the average nitrogen balance vary by farm type. 

Nitrogen from fertiliser makes the largest contribution to nitrogen input for general cropping and cereal farm types. Following the large decrease in fertiliser inputs in 2022-23, nitrogen balance on arable farm types rose considerably in 2023-24 as fertiliser input increased.  

For the average cereal farm, nitrogen balance rose by 250% to 34.2 kg N surplus/ha compared with the previous year. 

Nitrogen balance on the average general cropping farm rose dramatically (by 434%) in 2023-24 to 16 kg N surplus/ha. Increased high energy concentrate feed inputs and a fall in wheat outputs also contributed to higher average nitrogen balance for general cropping farms in 2023-24. Lower cereal yields can lead to nitrogen accumulation in the soil, driving an increase in nitrogen balance.  

Dairy farms, where the primary output is milk, have the highest average nitrogen balance of the farm types, estimated at 281.3 kg N surplus/ha in 2023-24. Nitrogen balance for the average dairy farm increased by 12% from the previous year. The rise in nitrogen balance on average dairy farms was driven by increased feed and fertiliser inputs.  

Average nitrogen balance increased for LFA cattle and sheep and LFA sheep farms in 2023-24. A fall in oilseed rape and barley outputs drove a 36% increase to 52.4 kg N surplus/ha on average LFA cattle and sheep farms. Average nitrogen balance rose by 76% to 38.1 kg N surplus/ha on LFA sheep farms, driven by increased fertiliser input.  

Lowland cattle and sheep and LFA cattle were the only farm types where the average nitrogen balance decreased in 2023-24. A decrease in fertiliser input and low energy concentrate feed inputs drove a 5% decrease to 56.3 kg N surplus/ha on average LFA cattle farms. For lowland cattle and sheep farms, nitrogen balance fell by 4% to 60.6 kg N surplus/ha. Increased fertiliser and low energy feed concentrates were offset by increased bean, draff and potato inputs. 

Mixed farms have a nitrogen balance comparable to livestock farms, with an estimated average of 43.4 kg N surplus/ha in 2023-24, an increase of 3% from the previous year. 

<div style="margin-top: 40px;"></div>

## Nitrogen use efficiency {.tabset}

### All farms

**Figure 6a: Average farm nitrogen use efficiency decreases**

Median farm nitrogen use efficiency, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7a, echo = FALSE, fig.height=5.5, fig.width=8}
fig7a

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/NUE_AllTypes.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/NUE_AllTypes.csv> <button class="btn pdf"> .csv </button> </a>

### Cereal

**Figure 6b: Average cereal farm nitrogen use efficiency decreases**

Median farm nitrogen use efficiency, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7b, echo = FALSE, fig.height=5.5, fig.width=8}
fig7b

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/NUE_Cereal.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/NUE_Cereal.csv> <button class="btn pdf"> .csv </button> </a>

### General cropping

**Figure 6c: Average general cropping farm nitrogen use efficiency decreases**

Median farm nitrogen use efficiency, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7c, echo = FALSE, fig.height=5.5, fig.width=8}
fig7c

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/NUE_GeneralCropping.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/NUE_GeneralCropping.csv> <button class="btn pdf"> .csv </button> </a>

### Dairy

**Figure 6d: Average dairy farm nitrogen use efficiency remains stable**

Median farm nitrogen use efficiency, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7d, echo = FALSE, fig.height=5.5, fig.width=8}
fig7d

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/NUE_Dairy.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/NUE_Dairy.csv> <button class="btn pdf"> .csv </button> </a>

### LFA sheep 

**Figure 6e: Average LFA (less favoured area) sheep farm nitrogen use efficiency decreases**

Median farm nitrogen use efficiency, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7e, echo = FALSE, fig.height=5.5, fig.width=8}
fig7e

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/NUE_LFASheep.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/NUE_LFASheep.csv> <button class="btn pdf"> .csv </button> </a>

### LFA cattle

**Figure 6f: Average  LFA (less favoured area) cattle farm nitrogen use efficiency decreases**

Median farm nitrogen use efficiency, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7f, echo = FALSE, fig.height=5.5, fig.width=8}
fig7f

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/NUE_LFACattle.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/NUE_LFACattle.csv> <button class="btn pdf"> .csv </button> </a>

### LFA cattle and sheep

**Figure 6g: Average  LFA (less favoured area) cattle and sheep farm nitrogen use efficiency decreases**

Median farm nitrogen use efficiency, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7g, echo = FALSE, fig.height=5.5, fig.width=8}
fig7g

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/NUE_LFACattleandSheep.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/NUE_LFACattleandSheep.csv> <button class="btn pdf"> .csv </button> </a>

### Lowland cattle and sheep

**Figure 7h: Average lowland cattle and sheep farm nitrogen use efficiency decreases**

Median farm nitrogen use efficiency, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7h, echo = FALSE, fig.height=5.5, fig.width=8}
fig7h

```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/NUE_LowlandCattleandSheep.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/NUE_LowlandCattleandSheep.csv> <button class="btn pdf"> .csv </button> </a>

### Mixed

**Figure 7i: Average mixed farm nitrogen use efficiency increases**

Median farm nitrogen use efficiency, 2019-20 to 2023-24. Shaded area shows the 95% confidence interval.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7i, echo = FALSE, fig.height= 5.5, fig.width= 8}
fig7i


```

<font size="3"><strong>Source:</strong> Scottish agriculture greenhouse gas emissions and nitrogen use 2023-24</font>

<b> Download data for this chart </b>

<a href = download/NUE_Mixed.xlsx> <button class="btn pdf"> .xlsx </button> </a> <a href = download/NUE_Mixed.csv> <button class="btn pdf"> .csv </button> </a>

##  {.unnumbered}

Nitrogen use efficiency, the ratio of nitrogen outputs to inputs, indicates the proportion of nitrogen used in the farm system in a given year. This is related to nitrogen balance and is a different way of measuring the relationship between outputs and inputs.

Nitrogen use efficiency was estimated at 29% for the average farm in 2023-24, a decrease of two percentage points compared with 2022-23 (31%). A higher NUE typically indicates a more efficient use of nitrogen. 

Estimates of nitrogen use efficiency vary by farm type. Farms engaged mostly in meat and milk production typically have much lower nitrogen use efficiency estimates than arable farms. Livestock based agriculture is inherently less nitrogen efficient than arable agriculture because only a small proportion of the ingested nitrogen by livestock ends up in useful nitrogen-containing produce. 

Within a farm type, farms can also vary widely in their nitrogen use efficiency due to management practices and environmental conditions. This suggests that there may be scope for some farms to increase nitrogen use efficiency.

For livestock farms, average nitrogen use efficiency was highest for lowland cattle and sheep in 2023–24, at 19%. This was a decrease of 9 percentage points compared with 2022-23. 

The average dairy farm had a nitrogen use efficiency of 19%, remaining stable since the previous year. 

Nitrogen use efficiency for the average cereal farm was 78%, falling by 12 percentage points from 2022-23. For the average general cropping farm, nitrogen use efficiency was 88%, a decrease of 11 percentage points.  

Mixed farms show an intermediate efficiency between arable and livestock farms and increased by two percentage points to 59% in 2023-24. 

The changes in nitrogen input and output between 2022-23 and 2023-24 that resulted in changes for both nitrogen balance and nitrogen use efficiency estimates, are discussed in the previous [section](#n-balance).

<br></br>

# About the statistics {#about-statistics}

<div style="margin-top: 30px;"></div>

## About GHG emissions in agriculture

There are three main GHGs produced in agriculture. These are carbon dioxide, methane, and nitrous oxide. Total emissions are presented in carbon dioxide equivalent (CO~2~e) units. This takes into account the different effects that the different gases have on climate change, known as their global warming potential (GWP). Over a 100-year period, methane's GWP is considered to be 28 times stronger than that of carbon dioxide, while nitrous oxide's GWP is considered to be 265 times stronger than carbon dioxide.

<div style="margin-top: 30px;"></div>

## About national GHG estimates for agriculture and agricultural subsectors {#table6}

The data source for the national estimates is the <a href ="https://www.gov.scot/collections/environment-statistics/" target ="_blank"> Scottish Greenhouse Gas Statistics</a>  where emissions are allocated by sector and type of gas. Emissions within the agriculture sector are grouped into the following categories:

-   agricultural combustion
-   agricultural soils
-   livestock
-   other agriculture - including lubricants and crop combustion.

This report groups emissions within the agriculture sector into subsectors. Subsector analysis assigns agricultural emissions to enterprise or activity type. This gives more detail about the emission sources within agriculture. This report provides emission estimates for the following subsectors:

-   suckler beef
-   dairy beef
-   dairy
-   sheep
-   arable
-   other - including pigs, other livestock, and non-agriculture.

**Table 6: Allocation of emission sources to agricultural subsector.**
  
Proportions are based on a <a href ="https://www.gov.scot/publications/disaggregating-headline-smart-inventory-figures/" target ="_blank">methodology developed by SRUC</a>. The sectoral split of emissions are based on proportions from 2018 and are applied to the entire timeseries.

<button class="btn btn-primary" data-toggle="collapse" data-target="#Table6">
  
Show/Hide data table

</button>



::: {#Table6 .collapse}
```{r table_6, echo = FALSE, results = 'markup'}
knitr::kable(table_6) %>% kableExtra::kable_styling(font_size = 16,  html_font = 'Arial', bootstrap_options= 'responsive')
caption = "Proportions are based on a methodology developed by SRUC. The sectoral split of emissions are based on proportions from 2018 and are applied to the entire timeseries."
```
:::
      
<strong> Download data for this table </strong>
      
::: {style="margin-bottom: 20px;"}
<a href = download/Allocation_to_subsector.xlsx> <button class="btn pdf"> .xlsx </button> </a><a href = download/Allocation_to_subsector.csv> <button class="btn pdf"> .csv </button></a>
:::
      

Further information about the data sources and methods are available in the <a href = methodology.html> Methodology</a>.

<div style="margin-top: 30px;"></div>

## About sources of agriculture GHG emissions

The possible sources of GHG emissions from agricultural activity are grouped from categories in the Scottish Greenhouse Gas statistics into:

-   **enteric fermentation** - a digestive process whereby carbohydrates are broken down by microorganisms into simple molecules. Both ruminant animals (e.g. cattle and sheep), and some non-ruminant animals (e.g. swine and horses) produce methane, although ruminants are the largest source per unit of feed intake.

-   **manure management** - describes the methods used to capture, store, treat and use animal manure within a farm system. The choice of manure management techniques can significantly alter the GHG emissions and wider environmental impact of the manure.

-   **agricultural soils**

-   **fuel combustion** - including emissions from static and mobile machinery.

-   **liming** - a common agricultural practice where lime is applied to soils to raise pH, optimising plant growth through improved nutrient uptake. Maintaining optimum pH has multiple benefits including minimising fertiliser waste and associated diffuse pollution.

-   **other emission sources** (including urea application, and non-energy products from fuels and solvent use). Urea-based fertilizers are applied to soils to improve the availability of nitrogen for plants. However increased nitrogen also impacts soil microbial activity, resulting in increased nitrous oxide emissions. Excess urea may leach into the environment causing diffuse pollution or volatilize into ammonia. Non-energy products from fuels and solvent use include lubricants, paraffin waxes, bitumen/asphalt, and solvents. Carbon dioxide is the major GHG released from the use of these products.

Some emission sources cut across several agriculture subsectors. For example, both dairy and arable farmers will produce emissions from the process of fuel combustion.

More information about how the emission sources have been grouped is available in the <a href = methodology.html#subsector> Methodology</a>.

<div style="margin-top: 20px;"></div>

## About nitrogen in agriculture

Nitrogen is an essential nutrient for plant growth. Nitrogen supplied by fertiliser, soil and other sources are consumed in harvested crops and animal products. However, excess nitrogen can also be lost to the wider environment. Losses of nitrogen to the environment can have harmful effects on, for example, climate change, biodiversity, air quality and water quality.

<div style="margin-top: 20px;"></div>

## About national nitrogen use estimates

National results for agriculture nitrogen use efficiency are available in the <a href ="https://www.gov.scot/publications/scottish-nitrogen-balance-sheet-2022/" target ="_blank">Scottish Nitrogen Balance Sheet</a>. Results are available for 2022.

<div style="margin-top: 20px;"></div>

## About average farm GHG emissions

Average farm estimates complement the national results by providing a view of emissions and nitrogen use at a whole-farm level and how these vary over time.

Farms are complex businesses with multiple enterprises that contribute to GHG emissions. For example, cereal farms may have livestock and the scale of this secondary enterprise can vary.

Average farm estimates include emissions from farm production activities. They do not account for any sequestration by woodland areas or soil.

Data do not cover the full agricultural industry and are not comparable with national estimates.

Farm data are collected through the Farm Business Survey. This provides a consistent and high-quality input data set about farm activities. It is a survey of around 400 commercial farms (whose standard output exceeds around £20,000 per year and are not part time). The sample excludes farm types not in receipt of support payments (such as pigs, poultry and horticulture). Results are for the accounting year 2022-23, which relates to the 2022 crop year. The survey sample is weighted using results from the 2022 June Agricultural Census.

Average results are shown for the median farm and as such are not directly comparable to results published on farm income.

Farm GHG emissions are estimated using a carbon footprint calculator tool (<a href ="https://www.agrecalc.com/" target ="_blank">Agrecalc</a>). It is based on the life cycle assessment (LCA) framework for estimating emissions from products and processes. The LCA accounts for emissions up to when product leaves the farm. The tool uses the latest IPCC Tier I and Tier II as well as national figures from the UK National GHG Inventory in its calculations. It is PAS2050:2011 certified. Different results may be obtained using different carbon calculators or different methods, for example using an inventory approach.

Results for 2021-22 onwards include methodological improvements to Agrecalc methodology for estimating GHG emissions. These estimates are produced using Agrecalc Cloud, whereas earlier estimates were produced using a legacy version of Agrecalc Web. Agrecalc Cloud methodology changes include:

-   updates to models based on IPCC (2019) guidelines

-   using Global Feed LCA Institute (GFLI) database for feed emissions

-   using Fertiliser Europe database for fertiliser emissions

-   new manure management and mitigation options

Previous years will not be revised as not all data are available. More information is available in the <a href = methodology.html> Methodology</a>.

<div style="margin-top: 20px;"></div>

### About average farm absolute emissions

Gross emission per hectare results include all enterprises on farm. Farms are complex businesses with multiple activities that contribute to GHG emissions. For example, cereal farms may have livestock, and the scale of this secondary enterprise can vary. 

Gross emission per hectare results are weighted to the June Agricultural Census population.

Gross emissions are emissions from a unit area of agricultural land (hectare) adjusted for forage quality. This accounts for large differences in rough grazing area and quality between farms.

<div style="margin-top: 20px;"></div>

### About average enterprise emission intensity

Emission intensity results are for the primary product of farm enterprises (activities). Enterprise level emission intensity estimates are not weighted because these are based on a sub-sample of results. These replace previously published whole-farm emission intensities.

For beef and sheep enterprises, emission intensities are calculated per kg of carcass deadweight (kg dwt). Milk emission intensities for dairy enterprises are calculated per kg of fat and protein corrected milk (kg FPC milk). Cereals emission intensities are calculated for total barley, wheat, oats and minor cereals enterprises (tonne crop).

Results for beef and sheep enterprises are based on assumptions about typical average weights of animals. This can lead to some higher or lower estimates where the sale weights of animals are higher or lower than average. 

Cereal emission intensities are based on the grouping of different cereal enterprises. Due to changes in the type of crop grown each year, some variation between yearly estimates may be attributable to changes in cereals composition. 

<div style="margin-top: 20px;"></div>

### About average farm nitrogen use estimates {#nitrogen}

Farm nitrogen estimates are based on standard estimates of nitrogen content in all inputs and outputs. The amount of nitrogen is estimated for all farm inputs and outputs up to farm gate, where possible. Limited information is available for some inputs and outputs.

Organic farms are excluded from these estimates as limited information is collected on their use of biological fixation of nitrogen. As such, the results are not directly comparable to other results based on the Farm Business Survey.

<div style="margin-top: 20px;"></div>

### About average farm nitrogen balance and use efficiency

Nitrogen use estimates in this report are:

-   Nitrogen balance (kg N surplus/ha)

-   Nitrogen use efficiency NUE (% N output/N input)

Nitrogen balance is the difference between total nitrogen input and output. A higher balance indicates less efficient use of nitrogen. It provides an estimate of the size of the nitrogen surplus not being captured in agricultural products that is potentially available for losses.

Nitrogen use efficiency (NUE) is the ratio of nitrogen outputs to inputs. It indicates the proportion of nitrogen used in the farm system. This measure allows for better comparison across farms. NUE values should always be interpreted in relation to nitrogen surpluses and nitrogen outputs.

A higher NUE typically indicates a more efficient use of nitrogen but very high values may indicate unsustainable "soil mining". The best range of NUE values depends on the type of farming activity as well as environmental conditions, livestock types and feed types.

<br></br>

# Future publications

Details of future publications can be found on our <a href ="https://www.gov.scot/publications/official-statistics-forthcoming-publications/" target ="_blank">forthcoming publications</a> page.

<br></br>
