---
title: "Scottish agriculture greenhouse gas and nitrogen use: 2022-23"
author: no
date: "27 June 2024"
output:
  html_document:
    template: _template.html
    toc: yes
    toc_float: yes
    toc_depth: 2
    self_contained: yes
sgtemplates:
  footer:
    link:
      - href: "accessibility.html"
        text: "Accessibility"
      - href: "contact.html"
        text: "Contact"
  header:
    site_branding: "Scottish agriculture greenhouse gas and nitrogen use: 2022-23"
    toc:
      sticky: false
  seo:
    description: "SEO (search engine optimisation) description"
    keywords: "agriculture 2022, agriculture emissions, scotland ghg, beef emissions"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
source("scripts/national_charts.R")
source("scripts/CN_analysis_build.R")

```

<a href = download/scottish_agriculture_ghg_n_use_data_tables.xlsx> <button class="btn pdf"> Download data in .xlsx format </button></a>

This report includes:

-   Estimates of greenhouse gas emissions for Scottish agriculture as reported in the [Scottish Greenhouse Gas (GHG) Statistics 2022](https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022/).

-   Estimates of greenhouse gas emissions by agricultural subsector. New analysis of the Scottish Greenhouse Gas Statistics 2022 allocates total GHG emissions from agriculture to agricultural subsectors. The subsectors align more closely with agricultural enterprises and provide a different breakdown to the categories used in the Scottish Greenhouse Gas Statistics.

-   Emission and nitrogen use estimates for the average farm based on commercial sized farms in the Farm Business Survey. Results include all enterprises on farm. Farms are complex businesses with multiple activities that contribute to GHG emissions. For example, cereal farms may have livestock and the scale of this secondary enterprise can vary.

-   Average farm estimates are for the accounting year 2022-23 rather than the 2022 calendar year. Methodology improvements have been made for emission estimates for the average farm for 2022-23. Previous years have not yet been revised. Average farm emission results for 2022-23 are not directly comparable to previous years.

# Key Points

## For the Agriculture sector

-   In 2022, greenhouse gas emissions from Scottish agriculture were at their lowest levels since the baseline period in 1990. Total emissions from agriculture in 2022 were 7.7 million tonnes of carbon dioxide equivalent (MtCO~2~e). This is a decrease of 3% (0.3 MtCO~2~e) compared with 2021.

-   Price increases of inorganic fertiliser in 2022 may have contributed to lower emissions from agriculture compared with 2021, as usage rates fell. Prior to the fall in 2022, agriculture emissions had been relatively stable. From 2016 to 2021, emissions were within the range 7.9 to 8.0 MtCO2e.

-   Total agriculture emissions have been in a long-term decline. Since 1990 emissions have fallen by 12% from 8.8 MtCO2e to 7.7 MtCO2e. Over the same period, total GHG emissions in Scotland have halved from 81.9 MtCO2e to 40.6 MtCO2e.

-    The long-term fall in emissions from agriculture was largely driven by falling livestock numbers. From 1990 to 2022 cattle and sheep numbers have decreased by 20% and 33%, respectively.

-   In 2022, emissions from arable and dairy subsectors both fell by 4% (0.1 MtCO2e and 0.05 MtCO2e respectively) compared to 2021. Suckler beef emissions fell by 3% (0.1 MtCO2e) from the previous year,

-    The suckler beef subsector is the largest contributor to total agriculture emissions. In 2022 emissions were 3.2 MtCO2e, 41% of agriculture emissions. Suckler beef emissions fell by 0.6 MtCO2e (17%) from 1990 to 2022. This was the largest absolute reduction across the agriculture subsectors. Emissions from sheep and dairy beef subsectors also fell by 17% from the baseline period. This follows the trend of declining livestock numbers over the same period.

-    Emissions from the arable subsector increased by 14% from 1990 to 2022. Arable farming accounts for most agriculture emissions from fuel combustion and liming. Emissions from fuel combustion and liming have increased by 42% and 38% respectively, over the same period.

## For the average farm

-   Farm-level emission results reflect the different nature of farm outputs from different farm types. A wide variation of results within each farm type suggests that there may be scope for some farms to decrease their emissions.

-   Dairy farms tend to have the highest emissions per hectare due to high stocking densities. In 2022-23, emissions from the average dairy farm were estimated at 17.8 tCO2e ha.

-   In 2022-23, emissions per hectare were higher for the average cereal farm (2.8 tCO2e/ha) than for the average general cropping farm (2.2 tCO2e/ha).

-   Emissions per unit of output, or emissions intensity, are typically higher for livestock farms than arable farms, reflecting the different nature of the outputs. In 2022-23, sheep farms in less favoured areas (LFA) had the highest average emissions intensity at 27.7 kg CO2e/kg output on average.

-   Dairy farms, where the primary output is milk, have a lower emission intensity than livestock farms. The average dairy farm had an emissions intensity of 1.5 kg CO2e/kg output in 2022-23.

-   Farm-level nitrogen use results show a wide variation within each farm type. Farms can vary widely in their nitrogen use efficiency due to management practices and environmental conditions. This suggests that there may be scope for some farms to improve their nitrogen use.

-   Nitrogen balance, the difference between nitrogen inputs and outputs, on the average Scottish farm was 38.2 kg N surplus/ha in 2022-23. This is a decrease of 28% from the previous year and may indicate that less nitrogen from agriculture was at risk of being lost to the environment. Most farm types show a decrease in nitrogen balance in 2022-23 as nitrogen outputs increased from the previous year.

-   Nitrogen use efficiency (NUE), the ratio of nitrogen outputs to inputs, was estimated at 31% for the average farm in 2022-23. This is an increase of five percentage points compared with 2021-22 (26%). A higher NUE typically indicates a more efficient use of nitrogen.

-   Similar trends in nitrogen balance and nitrogen use efficiency were seen for most farm types. In general, this was driven by increased farm output.

-   Large decreases in nitrogen balance, and corresponding increases in nitrogen use efficiency, were seen for general cropping (129%) and cereal (66%) farms compared with the previous year. A larger decrease in fertiliser input as well as increases in output drove these results for the average cereal farm.

# Introduction

## What you need to know

This report provides GHG emissions for the Scottish agriculture sector, agricultural subsectors and for average Scottish farms by farm type. Nitrogen use estimates are also included for average Scottish farms. Several data sources have been combined to provide a broad view of emissions and nitrogen use from the agriculture sector. More information is available in the about the statistics section.

Farm-level emission results are per hectare or per kilogram of output. This helps with comparing businesses of different sizes. Methodology differences mean average farm estimates are not directly comparable with national estimates. Methods differ in their coverage of the industry and emissions.

Estimates cover similar but different time periods. Emissions for the agriculture sector and its subsectors are based on a calendar year, whilst farm-level results are based on an accounting year.

It is important to note that agriculture and food production rely on natural processes. This means that agricultural activity will always cause a degree of GHG emissions. Emissions will also vary widely between different agricultural produce.

## An Official Statistics in Development Publication for Scotland

These statistics are official statistics in development. Official statistics in development may be new or existing statistics, and will be tested with users, in line with the standards of trustworthiness, quality, and value in the [Code of Practice for Statistics](https://code.statisticsauthority.gov.uk/the-code/).

Scottish Government statistics are regulated by the Office for Statistics Regulation (OSR). OSR sets the standards of trustworthiness, quality and value in the [Code of Practice for Statistics](https://code.statisticsauthority.gov.uk/the-code/) that all producers of official statistics should adhere to.

# Agriculture sector greenhouse gas emissions

#### Greenhouse gas emissions from agriculture at lowest levels

Total and agriculture greenhouse gas emissions in Scotland, 1990 to 2022

::: {.alert .alert-info}
Click and drag to zoom in on part of the chart. Double click to reset.
:::

```{r fig1, echo = FALSE}
fig1

```

<font size="2"><strong>Source:</strong> [Scottish Greenhouse Gas Statistics](https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022/) Data for this chart is taken from Table 1 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

GHG emissions from Scottish agriculture were 7.7 million tonnes of carbon dioxide equivalent (MtCO2e) in 2022. This is the lowest value since the baseline period in 1990, a decrease of 1.0 MtCO2e over this period.

Since the baseline period in 1990 (8.8 MtCO2e), total agriculture emissions have decreased by 12%. Total Scotland emissions have fallen by 50% over the same period, from 81.9 MtCO2e in 1990 to 40.6 MtCO2e in 2022. In 2022, Agriculture was the third largest source of emissions in Scotland behind Transport and Industry, using Climate Change Plan emission categories [LINK].

Compared with the previous year, total agriculture emissions by 3% (from 8.0 MtCO2e in 2021 to 7.7 MtCO2e in 2022). Increased fertiliser prices may have contributed to lower agriculture emissions in 2022, as usage rates fell. Results from the British Survey of Fertiliser Practice 2022 showed that the overall application rate for nitrogen on crops and grass in Scotland decreased by 19% compared to the previous year.

Prior to the fall in 2022, agriculture emissions had been relatively stable. From 2016 to 2021, emissions were within the range 7.9 to 8.0 MtCO2e.

More results for total Scotland GHG emissions, including for agriculture, are available in the [Scottish Greenhouse Gas Statistics](https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022).

# Agriculture sub-sector greenhouse gas emissions

#### Suckler beef has the largest share of Scottish agriculture greenhouse gas (GHG) emissions

Agriculture GHG emissions in Scotland, by subsector, 2022

::: {.alert .alert-info}
Click and drag to zoom in on part of the chart. Double click to reset.
:::

```{r fig2, echo = FALSE}
fig2


```

Source: [Scottish Greenhouse Gas Statistics](https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022/). Data for this chart is taken from Table 2 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)

The suckler beef subsector had the largest share of Scottish agriculture emissions from 1990 to 2022. In 2022 suckler beef contributed to 41% (3.2 MtCO2e) of total agricultural emissions. The second largest contribution is from arable (20% or 1.6 MtCO2e in 2022), closely followed by dairy (17% or 1.3 MtCO2e in 2022) and sheep (15% or 1.2 MtCO2e in 2022). Dairy beef and other subsectors (including pigs, poultry, and non-agriculture) accounted for a very small proportion (3%) of total agriculture emissions in 2022.

In 2022, emissions from arable and dairy subsectors fell by 4% compared with the previous year. Suckler beef emissions fell by 3% (0.1 MtCO2e) to 3.2 MtCO2e.

Reductions in emissions from suckler beef have contributed to the largest absolute decrease in total agriculture emissions since the baseline period. From 1990 to 2022 suckler beef emissions fell by 0.6 MtCO2e (17%). This was driven by declining cattle numbers over the same period. Cattle numbers fell from 2.1 million in 1990 to 1.7 million in 2022, a decrease of 20%.

The suckler beef subsector's contribution to total agriculture emissions has fallen from 43% in 1990 to 41% in 2022.

Emissions from arable farming, which has the next largest share of total emissions, increased by 14% from 1990 to 2022. This the only subsector with increased emissions compared with 1990. A rise in emissions from machinery and fertiliser use, both of which make large contributions to arable farming, drove the overall increase in the arable subsector emissions from 1990.

Emissions from both the sheep and dairy beef subsectors fell by 17% from 1990 to 2022.

## Sources of agriculture GHG emissions

#### Enteric fermentation is the largest source of agriculture emissions

Agriculture greenhouse gas emissions in Scotland, by subsector and emission source, 2022

::: {.alert .alert-info}
Click and drag to zoom in on part of the chart. Double click to reset. Click on individual emission sources on the right to select or deselect.
:::

```{r fig3, echo = FALSE}
fig3

```

Source: [Scottish Greenhouse Gas Statistics](https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022/). Data for this chart is taken from Table 3 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)

As reported in the Scottish Greenhouse Gas Statistics 2022, enteric fermentation (animal digestion) was the largest source of GHG emissions from agriculture in 2022. Emissions from enteric fermentation were 4.0 MtCO2e, just over half (52%) of total emissions from Scottish agriculture.

Looking at emission sources by subsector, just over half (52%) of emissions from enteric fermentation were produced by suckler beef.

Emissions from enteric fermentation have had the largest overall reduction compared with all other agricultural emission sources. From 1990 and 2022, emissions from enteric fermentation fell by 0.8 MtCO2e (17%). The primary producers of emissions from enteric fermentation are cattle and sheep. Over this period, the number of cattle and sheep in Scotland reduced by 20% and 33%, respectively.

Emissions from liming and fuel combustion have both increased from 1990 to 2022 by a combined 0.4 MtCO2e. Arable farming accounts for most emissions from liming and fuel combustion.

Another major source of emissions from agriculture is manure management, which is mostly associated with livestock.

# Farm GHG emissions

## Absolute gross emissions per hectare {.tabset}

### All farms

**Average farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5a, echo = FALSE, fig.height= 6}
fig5a
```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 4 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx).</font>

### Cereal

**Average cereal farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5b, echo = FALSE, fig.height= 6}
fig5b
```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 4 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx).</font>

### General cropping

**Average general cropping farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5c, echo = FALSE, fig.height= 6}
fig5c
```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 4 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx).</font>

### Dairy

**Average dairy farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5d, echo = FALSE, fig.height= 6}
fig5d
```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 4 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx).</font>

### Specialist sheep (LFA)

**Average specialist sheep LFA (less favoured area) farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5e, echo = FALSE, fig.height= 6}
fig5e
```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 4 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx).</font>

### Specialist cattle (LFA)

**Average specialist cattle LFA farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5f, echo = FALSE, fig.height= 6}
fig5f
```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 4 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx).</font>

### Cattle and sheep (LFA)

**Average cattle and sheep LFA farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5g, echo = FALSE, fig.height= 6}
fig5g
```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 4 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx).</font>

### Lowland cattle and sheep

**Average lowland cattle and sheep farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5h, echo = FALSE, fig.height= 6}
fig5h
```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 4 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx).</font>

### Mixed

**Average mixed farm absolute gross greenhouse gas (GHG) emissions**

Median farm absolute gross GHG emissions per hectare, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig5i, echo = FALSE, fig.height= 6}
fig5i
```

##  {.unnumbered}

::: {.alert .alert-info}
<strong>Guidance</strong> Zoom into plots using the magnifying glass. Hover over each box for the median, lower quartile and upper quartile.
:::

Absolute emissions for the average Scottish farm were estimated at 4.4 tCO2e/ha in 2022-23.

Emissions per hectare is a useful measure for arable farms where land area is more strongly related to output. In 2022-23, emissions for the average cereal farm were estimated at 2.8 tCO2e/ha. For general cropping, this was 2.2 tCO2e/ha.

# Emissions for the average cereal farm were more stable over the previous three years than those for general cropping farms. General cropping farms are likely to have more livestock and more intensive crops, e.g. potatoes, and have a different relationship between land area, output, and emissions.

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 4 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx).</font>

##  {.unnumbered}

Absolute emissions for the average Scottish farm were estimated at 4.4 tCO~2~e/ha in 2022-23.

Emissions per hectare is a useful measure for arable farms where land area is more strongly related to output. In 2022-23, emissions for the average cereal farm were estimated at 2.8 tCO~2~e/ha. For general cropping, this was 2.2 tCO~2~e/ha. Emissions for the average cereal farm were more stable over the previous three years than those for general cropping farms. General cropping farms are likely to have more livestock and more intensive crops, e.g. potatoes, and have a different relationship between land area, output, and emissions.

Dairy farms tend to have higher emissions per hectare than other livestock farms. This is due to higher stocking densities. In 2022-23, average dairy emissions were 17.8 tCO~2~e/ha. For the average lowland cattle and sheep farm (the next highest emitter), this was 6.9 tCO~2~e/ha.

Mixed farms represent a mix of crop and livestock activities and exhibit average emissions in between arable and livestock farm types. In 2022-23, average mixed farm emissions were 5.2 tCO~2~e/ha.

# Mixed farms represent a mix of crop and livestock activities and exhibit average emissions in between arable and livestock farm types. In 2022-23, average mixed farm emissions were 5.2 tCO2e/ha.

## Emission intensity {.tabset}

### All farms

**Average farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6a, echo = FALSE, fig.height= 6}
fig6a

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 5 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Cereal

**Average cereal farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6b, echo = FALSE, fig.height= 6}
fig6b

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 5 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### General cropping

**Average general cropping farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6c, echo = FALSE, fig.height= 6}
fig6c

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 5 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Dairy

**Average dairy farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6d, echo = FALSE, fig.height= 6}
fig6d

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 5 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Specialist sheep (LFA)

**Average specialist sheep LFA (less favoured area) farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6e, echo = FALSE, fig.height= 6}
fig6e

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 5 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Specialist cattle (LFA)

**Average specialist cattle LFA farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6f, echo = FALSE, fig.height= 6}
fig6f

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 5 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Cattle and sheep (LFA)

**Average cattle and sheep LFA farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6g, echo = FALSE, fig.height= 6}
fig6g

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 5 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Lowland cattle and sheep

**Average lowland cattle and sheep farm greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6h, echo = FALSE, fig.height= 6}
fig6h

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 5 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Mixed

**Average mixed farm emissions greenhouse gas (GHG) emission intensity**

Median farm GHG emission intensity, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles. Due to methodology improvements, results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig6i, echo = FALSE, fig.height= 6}
fig6i

```

##  {.unnumbered}

::: {.alert .alert-info}
<strong>Guidance</strong> <br> Zoom into plots using the magnifying glass. Hover over each box for the median, lower quartile and upper quartile.
:::

Emission intensity for the average farm was 7.0 kg CO2e/kg output in 2022-23.

Farms vary widely in their emission intensity results, suggesting that there may be scope for some farms to decrease emission intensity.

Livestock farms have higher emission intensities than arable or mixed farms. In 2022-23, sheep farms in less favoured areas (LFA) had the highest average emission intensity at 27.7 kg CO2e/kg output.

# Dairy farms, where the primary output is milk, have lower emission intensities compared with other livestock farms. The average dairy farm had an emission intensity of 1.5 kg CO2e/kg output in 2022-23.

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 5 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

##  {.unnumbered}

Emission intensity for the average farm was 7.0 kg CO~2~e/kg output in 2022-23.

Farms vary widely in their emission intensity results within each farm type, suggesting that there may be scope for some farms to decrease emission intensity.

Livestock farms have higher emission intensities than arable or mixed farms. In 2022-23, sheep farms in less favoured areas (LFA) had the highest average emission intensity at 27.7 kg CO~2~e/kg output. Dairy farms, where the primary output is milk, have lower emission intensities compared with other livestock farms. The average dairy farm had an emission intensity of 1.5 kg CO~2~e/kg output in 2022-23.

Cereals and general cropping farms had an average emission intensity of 0.4 and 0.3 kg CO~2~e/kg output respectively, in 2022-23. In the previous three years, both types remained similarly stable. Mixed farms showed an average emission intensity higher than arable farms, at 1.4 kg CO~2~e/kg.

Cereals and general cropping farms had an average emission intensity of 0.4 and 0.3 kg CO2e/kg output respectively, in 2022-23. In the previous three years, both types remained similarly stable.

Mixed farms showed an average emission intensity higher than arable farms, at 1.4 kg CO2e/kg.

# Farm nitrogen use

## Nitrogen balance {.tabset}

### All farms

**Average farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7a, echo = FALSE, fig.height= 6}
fig7a

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 6 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Cereal

**Average cereal farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7b, echo = FALSE, fig.height= 6}
fig7b

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 6 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### General cropping

**Average general cropping farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7c, echo = FALSE, fig.height= 6}
fig7c

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 6 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Dairy

**Average dairy farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7d, echo = FALSE, fig.height= 6}
fig7d

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 6 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Specialist sheep (LFA)

**Average specialist sheep LFA (less favoured area) farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7e, echo = FALSE, fig.height= 6}
fig7e

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 6 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Specialist cattle (LFA)

**Average specialist cattle LFA farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7f, echo = FALSE, fig.height= 6}
fig7f

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 6 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Cattle and sheep (LFA)

**Average cattle and sheep LFA farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7g, echo = FALSE, fig.height= 6}
fig7g

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 6 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Lowland cattle and sheep

**Average lowland cattle and sheep farm nitrogen balance increases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7h, echo = FALSE, fig.height= 6}
fig7h

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 6 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Mixed farm

**Average mixed farm nitrogen balance decreases**

Median farm nitrogen balance, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig7i, echo = FALSE, fig.height= 6}
fig7i

```

##  {.unnumbered}

::: {.alert .alert-info}
<strong>Guidance</strong> Zoom into plots using the magnifying glass. Hover over each box for the median, lower quartile and upper quartile.
:::

Nitrogen balance on the average Scottish farm was 38.2 kg N surplus/ha in 2022-23, a decrease of 28% from 2021-22. A large nitrogen balance can indicate a higher risk of nitrogen being lost to the environment.

On average, nitrogen output increased as farm outputs increased in 2022-23. Nitrogen inputs decreased, mostly because of decreased fertiliser input.

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 6 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

##  {.unnumbered}

Nitrogen balance measures the difference between nitrogen input and nitrogen output, accounting for land area.

Nitrogen balance on the average Scottish farm was 38.2 kg N surplus/ha in 2022-23, a decrease of 28% from 2021-22. A large nitrogen balance can indicate a higher risk of nitrogen being lost to the environment. More information about nitrogen estimates is available in [LINK].

For the average farm, the decrease in nitrogen balance was mostly driven by increases in nitrogen output. Nitrogen output increased as farm output increased in 2022-23. Nitrogen inputs also decreased, mostly as a result of decreased fertiliser input. Increased fertiliser prices may have contributed to this decrease in 2022-23, as usage rates fell.

Estimates of the average nitrogen balance vary by farm type. Estimates of nitrogen balance on the average general cropping farm decreased dramatically (129%) in 2022-23 to -5.0 kg N surplus/ha. A negative nitrogen balance is associated with a nitrogen use efficiency (NUE) that is above 100%. Both occur when nitrogen outputs are larger than nitrogen inputs. This may indicate a true loss of nitrogen from the soil and/or may be due to limitations of the methods used to capture all nitrogen inputs.

The average cereal farm also showed a large decrease (66%) in nitrogen balance to 9.4 kg N surplus/ha. Large decreases in fertiliser input and increases in output in 2022-23 compared with the previous year drove the decreases in nitrogen balance.

Dairy farms have the highest average nitrogen balance of the farm types, estimated at 255.6 kg N surplus/ha in 2022-23. Nitrogen balance for the average dairy farm decreased 7% from the previous year.

Lowland cattle and sheep were the only farm type where the average nitrogen balance increased in 2022-23, by 5% to 59.8 kg N surplus/ha.

Across the remaining livestock types, the average nitrogen balance decreased in 2022-23, most notably in LFA cattle and sheep, which decreased by 39% to 38.7 kg N surplus/ha. In LFA sheep and LFA cattle and sheep farms, decreases in nitrogen inputs drove this nitrogen balance decrease. In LFA cattle, increases in nitrogen outputs drove the decrease.

Estimates of the average nitrogen balance vary by farm type.

Nitrogen from fertiliser makes the largest contribution to nitrogen input for general cropping and cereal farm types.

Estimates of nitrogen balance on the average general cropping farm decreased dramatically (by 129%) in 2022-23 to -5.0 kg N surplus/ha. The average cereal farm also showed a large decrease (66%) in nitrogen balance to 9.4 kg N surplus/ha. Large decreases in fertiliser input and increases in output in 2022-23 compared with the previous year drove the decreases in nitrogen balance.

A negative nitrogen balance is associated with a nitrogen use efficiency (NUE) that is above 100%. Both occur when nitrogen outputs are larger than nitrogen inputs. This may indicate a loss of nitrogen from the soil but cannot be separated from known limitations of the methodology. Limited information is available about some nitrogen inputs. More information is available in the Methodology [LINK]

Dairy farms, where the primary output is milk, have the highest average nitrogen balance of the farm types, estimated at 255.6 kg N surplus/ha in 2022-23. Nitrogen balance for the average dairy farm decreased 7% from the previous year.

Lowland cattle and sheep were the only farm type where the average nitrogen balance increased in 2022-23, by 5% to 59.8 kg N surplus/ha.

Across the remaining livestock types, the average nitrogen balance decreased in 2022-23. This was most notable for the average LFA cattle and sheep farm, with a decrease of 39% to 38.7 kg N surplus/ha. In LFA sheep and LFA cattle and sheep farms, decreases in nitrogen inputs drove the nitrogen balance decrease. In LFA cattle, increases in nitrogen outputs drove the decrease.

Mixed farms have a nitrogen balance comparable to livestock farms, with an estimated average of 41.4 kg N surplus/ha in 2022-23, a decrease of 29% from previous years.

Mixed farms have a nitrogen balance comparable to livestock farms, with an estimated average of 41.4 kg N surplus/ha in 2022-23, a decrease of 29% from previous years.

## Nitrogen use efficiency {.tabset}

### All farms

**Average farm nitrogen use efficiency increases**

### Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig8a, echo = FALSE, fig.height= 6}
fig8a

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 7 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Cereal

**Average cereal farm nitrogen use efficiency increases**

Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig8b, echo = FALSE, fig.height= 6}
fig8b

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 7 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### General cropping

**Average general cropping farm nitrogen use efficiency increases**

Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig8c, echo = FALSE, fig.height= 6}
fig8c

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 7 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Dairy

**Average dairy farm nitrogen use efficiency remains stable**

Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig8d, echo = FALSE, fig.height= 6}
fig8d

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 7 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

# Specialist sheep (LFA)

### Specialist sheep (LFA)

**Average specialist sheep LFA (less favoured area) farm nitrogen use efficiency increases**

Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig8e, echo = FALSE, fig.height= 6}
fig8e

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 7 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Specialist cattle (LFA)

**Average specialist cattle LFA farm nitrogen use efficiency increases**

# Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig8f, echo = FALSE, fig.height= 6}
fig8f

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 7 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Cattle and sheep (LFA)

**Average cattle and sheep LFA farm nitrogen use efficiency increases**

Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig8g, echo = FALSE, fig.height= 6}
fig8g

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 7 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Lowland cattle and sheep

**Average lowland cattle and sheep farm nitrogen use efficiency increases**

Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig8h, echo = FALSE, fig.height= 6}
fig8h

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 7 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

### Mixed

**Average mixed farm nitrogen use efficiency decreases**

Median farm nitrogen use efficiency, 2019-20 to 2022-23. Box plot shows the median and upper and lower quartiles.

::: {.alert .alert-info}
Double click the chart to zoom in. Double click again to reset.
:::

```{r fig8i, echo = FALSE, fig.height= 6}
fig8i

```

<font size="2"><strong>Source:</strong> Farm Business Survey 2022-23. Data for this chart is taken from Table 7 of the [data tables](download/scottish_agriculture_ghg_n_use_data_tables.xlsx)</font>

##  {.unnumbered}

Nitrogen use efficiency, the ratio of nitrogen outputs to inputs, indicates the proportion of nitrogen used in the farm system in a given year. This is related to nitrogen balance and is a different way of measuring the relationship between outputs and inputs.

Nitrogen use efficiency was estimated at 31% for the average farm in 2022-23, an increase of five percentage points compared with 2021-22. This indicates that nitrogen outputs exceeded nitrogen inputs by an average of 68%.

Estimates of nitrogen use efficiency vary by farm type. Farms engaged mostly in meat and milk production typically have much lower nitrogen use efficiency estimates than arable farms. Livestock based agriculture is inherently less nitrogen efficient than arable agriculture because only a small proportion of the ingested nitrogen by livestock ends up in useful nitrogen-containing produce.

Within a farm type, farms can also vary widely in their nitrogen use efficiency due to management practices and environmental conditions. This suggests that there may be scope for some farms to increase nitrogen use efficiency. For livestock farms, average nitrogen use efficiency was highest for lowland cattle and sheep in 2022-23, at 28%. This is an increase of three percentage points compared with 2021-22.

The average dairy farm had nitrogen use efficiency of 19%, remaining stable since the previous year.

Nitrogen use efficiency for the average cereal farm was 90%, increasing by 10 percentage points in 2022-23. For the average general cropping farm, this was 102%, an increase of 15 percentage points.

Mixed farms show an intermediate efficiency between arable and livestock farms and decreased by one percentage point to 56% in 2022-23.

The changes in nitrogen input and output between 2021-22 and 2022-23 that resulted in changes for both nitrogen balance and nitrogen use efficiency estimates are discussed in the previous section.

::: {.alert .alert-info}
<strong>Guidance</strong> Zoom into plots using the magnifying glass. Hover over each box for the median, lower quartile and upper quartile.
:::

# Glossary

<b> Absolute gross GHG emissions (tonnes CO2 equivalent per hectare):</b> measures the total gross GHG emissions from a unit area of agriculture land (hectare).

<b>Carbon dioxide equivalent (CO2e):</b> provides an estimate of total GHG emission taking into account the different effects that different gases have on climate change, known as their global warming potential (GWP).

<b>Emissions intensity (kg CO2 equivalent per kg of output):</b> measures the carbon footprint of a unit of product up to the farm gate. It includes all types of farm product.

<b>Enteric fermentation:</b> a digestive process whereby carbohydrates are broken down by microorganisms into simple molecules. Both ruminant animals (e.g. cattle and sheep), and some non-ruminant animals (e.g. swine and horses) produce methane, although ruminants are the largest source per unit of feed intake.

<b>Farm type:</b> farms are classified based on the how much of their standard output is from the crop and livestock enterprises on each farm.

<b>Global warming potential (GWP):</b> estimates the relative climate change impact of different GHGs compared with carbon dioxide.

<b>Less Favoured Area (LFA):</b> land where farming is more difficult due to natural constraints, such as hills and soil quality.

<b>Liming:</b> a common agricultural practice where lime is applied to soils to raise pH, optimising plant growth through improved nutrient uptake. Maintaining optimum pH has multiple benefits including minimising fertiliser waste and associated diffuse pollution.

<b>Manure management:</b> describes the methods used to capture, store, treat and use animal manure within a farm system. The choice of manure management techniques can significantly alter the GHG emissions and wider environmental impact of the manure.

<b>Nitrogen balance (kg N surplus per hectare):</b> the difference between total nitrogen nput and output.

<b>Nitrogen use efficiency (% N output / N input):</b> the ratio of nitrogen outputs to inputs.

<b>Non-energy products from fuels and solvent use:</b> These include lubricants, paraffin waxes, bitumen/asphalt, and solvents. Carbon dioxide is the major GHG released from the use of these products.

<b>Standard output:</b> estimated worth of crops and livestock.

<b>Urea application:</b> Urea-based fertilizers are applied to soils to improve the availability of nitrogen for plants. However increased nitrogen also impacts soil microbial activity, resulting in increased nitrous oxide emissions. Excess urea may leach into the environment causing diffuse pollution or volatilize into ammonia.

# Data and Methodology

## How to access background or source data

The data collected for this publication are available...

## Methodology

More detailed information is available in the Methodology [link].

# Tell us what you think

We are always interested to hear from our users about how our statistics are used, and how they can be improved.

## Feedback survey

We'd appreciate it if you would complete our short feedback survey [LINK] on this publication.

## Enquiries

For enquiries about this publication please contact:

Jackie Massaya

Rural and Environmental Science and Analytical Services Statistics

e-mail: [agric.stats\@gov.scot](mailto:agric.stats@gov.scot){.email}

For general enquiries about Scottish Government statistics please contact:

Office of the Chief Statistician

e-mail: [statistics.enquiries\@gov.scot](mailto:statistics.enquiries@gov.scot){.email}

## Join our mailing list

If you would like to receive notification about statistical publications, or find out about consultations on our statistics please join the [ScotStat mailing list](http://www.gov.scot/scotstat).

## Future publications

Details of future publications can be found on our [forthcoming publications](https://www.gov.scot/publications/official-statistics-forthcoming-publications/) page.
