---
title: "Methodology"
date: "27 June 2024"
sgtemplates:
  footer:
    link:
    - href: accessibility.html
      text: Accessibility
    - href: contact.html
      text: Contact
  header:
    site_branding: "Scottish agriculture greenhouse gas emissions and nitrogen use:
      2022-23"
    toc:
      sticky: no
  navigation:
    breadcrumb_trail:
    - href: index.html
      text: Home
    - text: Methodology
    toc:
      sticky: no
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("scripts/data_sort.R")

```

# Introduction

This report provides information on the methodology for estimating greenhouse gas (GHG) emissions for Scottish agricultural subsectors and GHG emission and nitrogen use estimates for average Scottish farms by farm type. Information is included about estimation methods and limitations.

Estimates of GHG emissions for the Scottish agriculture sector are as reported in the Scottish Greenhouse Gas Statistics 2022. Methodology for these results <a href = "https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022/pages/section-d--revisions-to-the-inventory-and-methodology/" target ="_blank">is available alongside the publication</a>.

Estimates of GHG emissions by agricultural subsector are created by allocating total GHG emissions, as reported in the Greenhouse Gas Statistics 2022, to agricultural subsectors.

Average farm greenhouse gas and nitrogen use estimates are produced using data collected in the Farm Business Survey. Methodology and quality information about data collected in the FBS and its headline measure of income, Farm Business Income (FBI), is available in the <a href = "https://www.gov.scot/publications/scottish-farm-business-income-annual-estimates-methodology/" target ="_blank">Scottish farm business income: annual estimates: methodology</a>.

Methodology improvements have been made for 2022-23 emission estimates for the average farm in the 'Scottish agriculture greenhouse gas emissions and nitrogen use: 2022-2023' publication. Results for 2022-23 are not directly comparable to previous years. Previous years have not yet been revised. More information available under [Changes to methodology](#changes).

Details about greenhouse gas (GHG) emissions and nitrogen use estimates are included here in line with the <a href = "https://ec.europa.eu/eurostat/web/quality/european-quality-standards/quality-assurance-framework" target ="_blank">European Statistical System (ESS) quality framework</a>. This covers areas of statistical:

-   Relevance

-   Accuracy and reliability

-   Timeliness and punctuality

-   Accessibility and clarity

-   Coherence and comparability

## An official statistics in development publication for Scotland

These statistics are official statistics in development. Official statistics in development may be new or existing statistics, and will be tested with users, in line with the standards of trustworthiness, quality, and value in the Code of Practice for Statistics.

We wish to involve users in our assessment of suitability and quality. If you use this data we would like to hear from you. Please give us your feedback in this short survey, or alternatively please get in touch with us at [agric.stats@gov.scot](mailto:agric.stats@gov.scot).

## Developing these statistics

Farm level GHG emissions and nitrogen use estimates were first published in June 2023 as <a href = "https://www.gov.scot/publications/farm-business-survey-2021-22-farm-level-emissions-and-nitrogen-usage/" target ="_blank">Farm Business Survey 2021-22: Farm level emissions and nitrogen usage.</a>

This latest publication includes national GHG emissions of agriculture, to give a broader picture of the agricultural sector. New subsector analysis assigns agricultural emissions to enterprise or activity type. This gives more detail about the emission sources within agriculture. The name change of the publication reflects this additional content.

Scottish Government statistics are regulated by the Office for Statistics Regulation (OSR). OSR sets the standards of trustworthiness, quality and value in the <a href = "https://code.statisticsauthority.gov.uk/the-code/" target ="_blank">Code of Practice for Statistics</a> that all producers of official statistics should adhere to.

<br></br>


# Changes to methodology {#changes}

Methodology improvements have been made for 2022-23 farm-level GHG emission estimates. This means that 2022-23 absolute emissions and emission intensity estimates are not directly comparable with data from previous years.

Farm GHG emissions are produced using the Agricultural Resource Efficiency Calculator <a href = "https://www.agrecalc.com/" target ="_blank">(Agrecalc)</a>. For 2022-23, the following changes were made:

- Calculations: Fertiliser calculations use UK National Inventory 1990-2019 methodology

- Coefficients: IPCC 5th Assessment Report (AR5) GWP 100 conversion factors. Coefficients for fertiliser and organic manure application are sourced from the UK National Inventory 1990-2019 methodology

Agrecalc have published more information about <a href = "https://www.agrecalc.com/home/insights/how-will-my-carbon-footprint-change-in-agrecalc-cloud/" target ="_blank">changes to the platform</a>.

<br></br>

# Agricultural subsector estimates methodology {#subsector}

Estimates of GHG emissions by agricultural subsector are created by allocating total GHG emissions, as reported in the <a href = "https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022/" target ="_blank">Scottish Greenhouse Gas Statistics 2022</a> to agricultural subsectors.

The subsectors align more closely with agricultural enterprises and provide a different breakdown to the categories used in the Scottish Greenhouse Gas Statistics. Agriculture emission estimates are allocated to the following subsectors:

-   suckler beef

-   dairy beef

-   dairy

-   sheep

-   arable

-   other -- including pigs, other livestock, and non-agriculture.

Table 1 gives the proportions used to allocate emission sources to agricultural subsectors.

**Table 1: Allocation of emission sources to agricultural subsector.**

Proportions are based on a <a href ="https://www.gov.scot/publications/disaggregating-headline-smart-inventory-figures/" target ="_blank">methodology developed by SRUC</a>. The sectoral split of emissions are based on proportions from 2018 and are applied to the entire timeseries.

<button class="btn btn-primary" data-toggle="collapse" data-target="#Table1">

Show/Hide data table

</button>

::: {#Table1 .collapse}
```{r table_1_pct, echo = FALSE, results = 'markup'}
knitr::kable(table_1_pct) %>% kable_styling(font_size = 16,  html_font = 'Arial', bootstrap_options= 'responsive')
caption = "Proportions are based on a methodology developed by SRUC. The sectoral split of emissions are based on proportions from 2018 and are applied to the entire timeseries."
```
:::

<strong> Download data for this table </strong>

::: {style="margin-bottom: 20px;"}
<a href = download/Table_1.xlsx> <button class="btn pdf"> .xlsx </button> </a><a href = download/Table_1.csv> <button class="btn pdf"> .csv </button></a>
:::

The subsector split of emissions are based on proportions from 2018 which are applied to the entire timeseries. This approach does not take into account year-on year changes in the underlying data. This is a known limitation that impacts the accuracy of subsector estimates. Results should be considered as indicative rather than definitive.

## Emission sources

The possible sources of GHG emissions from agricultural activity are grouped from categories in the Scottish Greenhouse Gas statistics into:

-   enteric fermentation

-   manure management

-   fuel combustion

-   liming

-   other emission sources - including urea application, and non-energy products from fuels and solvent use.

Table 2 gives the groupings of IPCC emission source categories into these source categories.

**Table 2 Allocation of IPCC emission source categories to source categories in the Scottish agriculture GHG emissions and nitrogen use publication**

<button class="btn btn-primary" data-toggle="collapse" data-target="#Table2">

Show/Hide data table

</button>

::: {#Table2 .collapse}
```{r table_2_merge, echo = FALSE, results = 'markup'}
knitr::kable(table_2_merge) %>% kable_styling(font_size = 16,  html_font = 'Arial', bootstrap_options= 'responsive') %>% column_spec(1, bold = TRUE)
```
:::

<strong> Download data for this table </strong>

::: {style="margin-bottom: 20px;"}
<a href = download/Table_2.xlsx> <button class="btn pdf"> .xlsx </button> </a><a href = download/Table_2.csv> <button class="btn pdf"> .csv </button></a>
:::

<br></br>

# The Farm Business Survey (FBS) {#the-farm-business-survey-fbs}

More detailed methodology and quality information about data collected in the FBS and its headline measure of income, namely Farm Business Income (FBI), is available in <a href = "https://www.gov.scot/collections/scottish-farm-business-income-fbi-annual-estimates/" target ="_blank">Scottish farm business income (FBI)</a>.

The FBS results are obtained from a sample of around 400 Scottish farms that are stratified by farm type and economic size.

An important feature of the survey is the measurement of changes over time for particular types of farm. To achieve this, it is necessary to maintain farms in the sample surveyed over a number of years. The survey is effectively a panel survey with little change in the membership of the sample between years.

The survey is not carried out on a calendar-year basis but based on individual businesses' accounting year ends which all centre on the same cropping period and on average roughly align with the financial accounting year. For example, 2022-23 data is for the 2022-23 accounting year and relates to the 2022 crop year.

## Industry coverage and limitations for emission estimates

The FBS is designed to provide good coverage of the majority of economic output from supported sectors of agriculture.

The FBS is representative of around 10,500 farms in Scotland, which in 2022-23 is around 19% of all farms included in the June Agricultural Census. These farms cover 63% of Scotland's agricultural land, account for 45% of the Scottish farming labour force, and produce 73% of total standard output.

The survey is designed for coverage of commercial sized farms, and represents 95% of standard output from these farm types. The survey is restricted to farms with considerable economic activity (at least €25,000 of output, equivalent to approximately £20,000) that are not considered spare-time farms (have a Standard Labour Requirement (SLR) of more than 0.5).

The survey does not include farms predominantly engaged in horticulture, poultry, egg production or pig production.

Data from the June Agricultural Census are used to weight the farm business income estimates. A calibrated weighting method sums the following variables to equal the estimated population totals:

-   type of farm

-   tenure type of farm

-   area of barley

-   area of oats

-   area of potatoes

-   area of wheat

-   rented area of land

-   total area of farm

-   number of dairy cows

-   number of beef cows

-   number of ewes

## Classification of Farms

The classification is based on detailed sub-types as defined in the <a href = "https://eur-lex.europa.eu/legal-content/EN/TXT/PDF/?uri=CELEX:32014R1198&from=en" target ="_blank">European Commission (EC) farm typology</a>, which have been grouped together where required to give the types shown below.

The classification is based on the relative importance of the various crop and livestock enterprises on each farm assessed in terms of standard output. The method of classifying each farm is to multiply the area of each crop (other than forage) and the average number of each category of livestock by the appropriate standard output, with the largest source of output determining the type of farm. The list below defines the main types that are reported in the FBS and groupings applied to present emissions and nitrogen use data.

**Specialist Sheep (LFA)** - Farms in less-favoured areas with more than two-thirds of their total standard output coming from sheep.

**Specialist Beef (LFA)** - Farms in less-favoured areas with more than two-thirds of their total standard output coming from cattle.

**Cattle and Sheep (LFA)** - Farms in less-favoured areas with more than two-thirds of their total standard output coming from sheep and beef cattle together.

**Cereals** - Farms with more than two-thirds of their total standard output coming from cereals and oilseeds.

**General Cropping** - Other farms with more than two-thirds of their total standard output coming from all crops.

**Dairy** - Farms with more than two-thirds of their total standard output coming from dairy cows.

**Lowland livestock** - Farms NOT in less-favoured areas, with more than two-thirds of their total standard output coming from sheep and beef cattle.

**Mixed** - Farms with no enterprise contributing more than two-thirds of their total standard output.

<br></br>

# Farm level GHG emission estimates methodology

The 'Scottish agriculture greenhouse gas emissions and nitrogen use' publication reports on the following headline metrics of GHG emissions:

-   Absolute gross GHG emissions (tCO~2~e/ha)

-   Emission intensity (CO~2~e/kg output)

Absolute gross emissions are emissions from a unit area of agricultural land (hectare) adjusted for forage quality. This accounts for large differences in rough grazing area and quality between farms.

Emission intensity is the carbon footprint of a unit of product or output (kg) up to the farm gate. It includes all types of agricultural products produced on the farm.

Data collected through the Farm Business Survey provide a consistent and high quality input data set about farm activities. GHG emissions are then estimated using a carbon footprint calculator tool. They do not account for any sequestration by woodland areas or soil. Transport of product or livestock off farm is not included.

Different results may be obtained using different carbon calculators or different methods, for example using an inventory approach.

The Agricultural Resource Efficiency Calculator <a href = "https://www.agrecalc.com/" target ="_blank"> Agrecalc</a> is used. This is a farm carbon footprint tool developed by the consulting division of Scotland's Rural College (SRUC). It can estimate greenhouse gas (GHG) emissions from agriculture and identify their main sources. It can also be used by farms to benchmark key performance indicators and simulate mitigation measures and what-if scenario for planning purposes.

The three main greenhouse gases produced in agriculture are estimated by Agrecalc:

-   carbon dioxide, produced by burning fossil fuels

-   methane, a natural by-product of animal digestion

-   nitrous oxide, released from soils following the application of nitrogen fertiliser (manufactured and organic) and soil disturbance.

Emissions are calculated for the whole farm and per unit (kg) of product. All mainstream agricultural enterprises are included. This includes cattle, sheep, dairy, pigs, poultry, cereals, oilseeds, potatoes, vegetables, and fruits.

Total emissions are presented in carbon dioxide equivalent (CO~2~e) units. This takes into account the different effects that the different gases have on climate change, known as their global warming potential (GWP). Over a 100-year period, methane's GWP is considered to be 28 times stronger than that of carbon dioxide, while nitrous oxide's GWP is considered to be 265 times stronger than carbon dioxide.

Agrecalc is based on the life cycle assessment (LCA) framework for estimating emissions from products and processes. The LCA accounts for emissions up to when product leaves the farm. It uses the latest Intergovernmental Panel on Climate Change (IPCC) Tier I and Tier II guidelines as well as national figures from the UK National Greenhouse Gas Inventory in its calculations. It is PAS2050:2011 certified.

The <a href ="https://www.ipcc.ch/report/2019-refinement-to-the-2006-ipcc-guidelines-for-national-greenhouse-gas-inventories/" target ="_blank"> IPCC methods for greenhouse gas reporting</a> are split into three tiers of increasing complexity and specificity. Tier I reporting standards use default figures published by the IPCC which provide a general estimate of greenhouse gas emissions, but likely miss important sources of variance. Tier II reporting standards are slightly more specific than Tier I, as they use national research to generate country-specific emission factors. Finally, Tier III reporting uses process-based models to predict emissions with the greatest accuracy and system-specificity.

To balance model performance and data requirements, Agrecalc makes use of higher Tier methods for large emissions sources. Default methods are used where higher Tier methods would increase data requirements beyond what is generally available on a farm. They are also used for smaller emission sources, and for emission sources where more research is needed to improve the resolution of emission factors.

In the next sections, a description of how IPCC methodology is used in each enterprise is provided.

### Beef and dairy

The beef and dairy models are based on IPCC Tier II guidelines. This Tier II calculation includes detailed modelling of the energy requirements of beef and dairy herds based on activity levels, growth rates, life stages, gender, and climate. Enteric methane emissions and emissions from manure deposited on grazing lands are also calculated using IPCC Tier II methods published in the most recent guidance.

### Sheep

The sheep model is also based on the latest IPCC Tier II guidelines. This Tier II calculation includes detailed modelling of the energy requirements of sheep flocks based on activity levels, growth rates, life stages, gender, and climate. Like the beef and dairy models, this model estimates emissions from enteric fermentation and manure deposited on grazing lands using IPCC Tier II methodology.

### Pigs

The pig model goes beyond IPCC reporting standards, incorporating a detailed energy balance model for pigs published by Food and Agriculture Organisation (FAO). This expansion upon standard methods allows the model to use Tier II reporting standards for pigs, improving the detail of the pig model outputs beyond that provided by standard greenhouse gas reporting for pigs.

### Poultry

The poultry model also goes beyond IPCC reporting standards, incorporating a detailed energy balance model for poultry published by FAO. The outputs of this model are in line with IPCC Tier II methods. This expansion upon standard methods allows the model to estimate emissions at a similar level of detail to our beef and dairy models.

### Manure Management

The manure management model estimates emissions related to storage and treatment of manures. The model uses Tier II methods to calculate methane emissions from manure, which uses the dietary characteristics of livestock to calculate methane emissions. For liquid storage systems, the model directly interacts with climate data to estimate methane emissions, in line with IPCC Tier II guidelines. The model also calculates nitrous oxide emissions from manures using Tier II methods, which incorporate information about the nitrogen content of livestock diets.

### Arable Enterprises and Improved Grassland

The arable farming model is based on IPCC Tier I and Tier II factors for cropland management. This includes Tier II factors for direct nitrous oxide emissions from organic and inorganic fertilisers, derived from the UK National Inventory. Tier I emission factors are utilised for indirect nitrous oxide emissions related to volatisation and leaching. Nitrous oxide emissions from crop residues are also calculated using IPCC Tier I methods.

### Embedded Resource Emissions

The carbon footprint model draws on various external databases to estimate emissions from imported feed, fuel, electricity, and fertiliser inputs. For emissions related to energy use, the tool uses figures published by Defra for UK energy use and figures published by the Greenhouse Gas Protocol (2012) for energy use in other countries. For the embedded emissions of fertilisers, the tool draws on a study by Fertilisers Europe (2018) which reported emissions related to different fertilisers by region. For emissions related to imported feed rations, the tool currently uses the Dutch Feedprint database and are in the process up updating these values to the Global Feed LCA Institute (GFLI). However, work on migrating to the GFLI database is currently undertaken to allow the users to choose from a greater variety of feed ingredients.

<br></br>

# Farm level nitrogen use estimates methodology

The 'Scottish agriculture greenhouse gas emissions and nitrogen use' publication reports on the following headline metrics of nitrogen use:

-   Nitrogen balance (kg N surplus/ha)

-   Nitrogen use efficiency (NUE, %)

Nitrogen balance is the difference between total nitrogen input and ouput. A higher balance indicates less efficient use of nitrogen. It provides an estimate of the size of the nitrogen surplus not being captured in agricultural products that is potentially available for losses.

NUE is the ratio of nitrogen outputs to inputs. It indicates the proportion of nitrogen used in the farm system. This measure allows for better comparison across farms. NUE values should always be interpreted in relation to nitrogen surpluses and nitrogen outputs.

A higher NUE typically indicates a more efficient use of nitrogen but very high values may indicate unsustainable "soil mining". The best range of NUE values depends on the type of farming activity as well as environmental conditions, livestock types and feed types.

Estimate methodology follows guidance from the <a href = "https://www.researchgate.net/publication/312554339_Nitrogen_Use_Efficiency_NUE_-_an_indicator_for_the_utilization_of_nitrogen_in_agriculture_and_food_systems_Prepared_by_the_EU_Nitrogen_Expert_Panel" target ="_blank">EU Nitrogen Expert Panel</a>  for assessing nitrogen at farm level.

This method assumes standard quantities for the nitrogen content of inputs and outputs for each year. The amount of nitrogen is estimated for all farm inputs and outputs up to farm gate, where possible.

Total nitrogen input is estimated as the nitrogen in operating resources and feed ( in kg N per hectare per year). It includes estimates for nitrogen in:

-   fertiliser

-   imported feed and fodder

-   biological fixation from peas and beans

-   atmospheric deposition

-   seed and planting material

-   bedding material

-   imported manure

-   irrigation water

-   animal manure, compost and sewage sludge

Limited information is available about the use of clover for biological fixation of nitrogen. As such, organic farms are excluded from estimates. Organic farms make up around 5% of the Farm Business Survey sample.

Limited information is available about farm grown feed and fodder and no estimate is included.

Seed quantity for the purpose of estimating a nitrogen input is calculated by assuming sowing rates follow best practice for crops.

Limited information is available about type of fertiliser, farm stocks of slurry, farmyard manure and compost, or application methods. This limits the quality of the nitrogen input estimates and understanding of the potential for nitrogen leaching. Manure estimates are based on the average number of animals on farm over the year and no attempt is made to account for changes in manure stocks.

Total nitrogen output is estimated as the nitrogen in produce exported from farm (kg N / farmed area (ha)) (kg N/ha/yr). It includes estimates for nitrogen in:

-   crop products

-   livestock sold

-   livestock products

The method does not take into account the changing status of nitrogen in soils over time.

Crops products, including fodder, are included where these are for sale. There are occasionally large sales of fodder which result in high NUE estimates for some farms.

<br></br>

# Relevance

Relevance is about making sure our statistics meet the needs of users.

These data are designated as official statistics in development. They are newly developed statistics undergoing testing.

Agriculture and food production rely on natural processes and consequently will always cause a degree of greenhouse gas emissions, the primary cause of global warming. A high proportion of Scottish emissions are from agriculture.

## Agriculture sector and subsector estimates

This latest publication includes national GHG emissions of agriculture, to give a broader picture of the agricultural sector.

New subsector analysis assigns agricultural emissions to enterprise or activity type. This gives more detail about the emission sources within agriculture. Stakeholder engagement identified a need for the breakdown of agriculture emissions into relevant enterprises.

## Farm level GHG emissions and nitrogen use estimates

This report provides greenhouse gas emission estimates and nitrogen use estimates for the average Scottish farm over time. These data complement national level estimates and enterprise specific estimates by providing a view and timeseries of emissions and nitrogen use on farm.

Nitrogen is a key driver of productivity in agriculture given its direct impact on yield. Nitrogen fertiliser is an expensive input and efficient use is linked to profitability. Nitrogen can also harm the environment and is linked to water pollution, poorer air quality, climate change and damage to natural ecosystems.

The emission and nitrogen use indicators published in 'Scottish agriculture greenhouse gas emissions and nitrogen use' were selected following a <a href = "https://sefari.scot/environmental-indicators-in-farm-level-surveys-a-literature-review" target = "_blank">systematic literature review </a> of environmental indicators and metrics that are currently used in farm level surveys and <a href = "https://sefari.scot/environmental-indicators-in-farm-level-surveys-stakeholder-engagement" target = "_blank">user engagement</a> undertaken in 2023 and 2024.

Findings from stakeholder engagement showed that the most common indicators found in our literature review (absolute emissions and emission intensity as well as nitrogen use efficiency and nitrogen surplus) also most commonly meet the needs of our users.

Other indicators can be estimated from these data and may be available on demand and considered for future publications. User need was also identified for indicators that that cannot currently be produced as limited data is collected.

The data are under development to continue to improve and understand data quality and ensure that analysis is fit for purpose.

We wish to involve users in our assessment of suitability and quality. If you use this data we would like to hear from you, please get in touch with us at [agric.stats@gov.scot](mailto:agric.stats@gov.scot).

<br></br>

# Accuracy and reliability

This section discusses how accurately and reliably these statistics portray reality.

These data are designated as official statistics in development as they are still in development, with a potentially wider degree of uncertainty in the resulting estimates as the methods and processes are established and verified.

## Scottish Greenhouse Gas Statistics

Detail about revisions to the GHG Inventory and methodology are available at <a href = "https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022/pages/section-d--revisions-to-the-inventory-and-methodology/" target = "_blank">Scottish Greenhouse Gas Statistics 2022</a>.

## Agriculture subsector estimates

There are limitations to the agriculture subsector emissions estimates methodology. More information is included under [Agricultural subsector estimates methodology](#subsector)

## June Agricultural Census area estimates

Farm level emission estimates are weighted using the same method as farm business income, which relies on results from the June Agricultural Census. There may be limitations in applying a weighting methodology that was developed for economic outputs to emissions results. More information is under [The Farm Business Survey (FBS)](#the-farm-business-survey-fbs).

Data quality and assurance measures for the June Agricultural Census are available at <a href = "https://www.gov.scot/collections/june-scottish-agricultural-census/" target = "_blank">Scottish Agricultural Census: results</a>.

## The Farm Business Survey

Data quality and assurance measures for the Farm Business Survey and its official statistical outputs are available at <a href = "https://www.gov.scot/collections/scottish-farm-business-income-fbi-annual-estimates/" target = "_blank">Scottish farm business income (FBI)</a>.

As a sample survey, results are subject to a degree of uncertainty. More detailed breakdowns of the sample result in relatively low sample sizes and an increase in uncertainty. For example in representing overall national averages by farm type.

## Farm level GHG emission and nitrogen use estimates

To demonstrate the variability in the data the average (median) result is presented along with the upper and lower quartiles of the data. Results are not directly comparable to results published on farm income for the average (mean) farm.

There are limitations to the farm level emissions estimate methodology. Different results may be obtained using different carbon calculators or different methods, for example using an inventory approach. More information is included under [Farm level GHG emission estimates methodology].

There are limitations to the farm level nitrogen estimate methodology. Nitrogen estimates are based on standard estimates of nitrogen content in all farm inputs and outputs where possible. Nitrogen estimates are not made for organic farms, which means a small proportion of the sample are excluded. More information is included under [Farm level nitrogen use estimates methodology].

<br></br>

# Timeliness and punctuality

Timeliness is about releasing statistics in a timely and punctual manner.

The 'Scottish agriculture greenhouse gas emissions and nitrogen use: 2022-2023' publication was first released on 27 June 2024 at <a href = "https://data.gov.scot/scottish-agriculture-greenhouse-gas-emissions-nitrogen-use-2022-23/" target ="_blank">Scottish agriculture greenhouse gas emissions and nitrogen use: 2022-23</a>.

## Agriculture subsector estimates

Agriculture emissions are reported in <a href="https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022/" target = "_blank">Scottish Greenhouse Gas Statistics 2022</a>, which was released on 18 June 2024. Subsector analysis of agriculture emissions cannot be published earlier than Scottish Greenhouse Gas Statistics. This publication was released shortly after, allowing time for quality assurance.

## Farm level GHG emissions and nitrogen use estimates

Farms in the Farm Business Survey have accounting year-ends between November and May. Data collection for farms with May year-ends cannot begin until June. Data collection, fieldwork and processing takes several months, and the finalised dataset is passed to the Scottish Government in December.

Economic results from the survey are released in March of the following year. This allows sufficient time for data processing, analysis and quality assurance, as well as compilation, preparation and dissemination of final results.

Results on farm level emissions and nitrogen usage are processed after completion of the economic results that contribute to the National Statistics Publication on Farm Business Income and cannot be released earlier than March of the year following the data collection period.

<br></br>

# Accessibility and clarity

Accessibility and clarity are about:

-   the presentation of statistics in a clear and understandable form

-   the release of statistics in a suitable and convenient manner

-   availability and accessibility on an impartial basis with

-   availability and accessibility of supporting metadata and guidance

These statistics are published online by the Scottish Government in an accessible format (html).

Data tables are made available in excel format to allow users to carry out further analysis. Charts in the report are available to download in excel and csv format. Farm-level datasets may be made available to recognised research organisations for research purposes agreed with the Scottish Government with appropriate security and disclosure control in place. To make a data request get in touch with us at [agric.stats@gov.scot](mailto:agric.stats@gov.scot).

Methodological information and specific quality issues are included on these pages. Metadata appropriate for users to consider comparability of statistics over time is included with all data releases, for example as notes on tables. This metadata is regularly reviewed and updated as necessary with each publication or more frequently as needed.

We aim to use modern and accessible means of dissemination and communication of statistics and are committed to continual improvement. More information about general accessibility is available under  [Accessibility](https://data.gov.scot/scottish-agriculture-greenhouse-gas-emissions-nitrogen-use-2022-23/accessibility.html)

<br></br>

# Coherence and comparability

This section covers how consistent these statistics are over time and how comparable they are with those of other regions and countries.

## Agriculture subsector estimates

Detail about revisions to the GHG Inventory and methodology are available at <a href = "https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022/pages/section-d--revisions-to-the-inventory-and-methodology/" target = "_blank">Scottish Greenhouse Gas Statistics 2022 </a>.

This is the first release of the timeseries for subsector estimates. Results are available from 1990 to 2022.

## Farm level GHG emissions and nitrogen use estimates

Methodology improvements have been made for 2022-23 farm-level GHG emission estimates. This means that 2022-23 absolute emissions and emission intensity estimates are not directly comparable with data from previous years. More information is included under [Changes to methodology]. Results from 2019-20 until 2021-22 are considered to be consistent and comparable over time.

Nitrogen use results are considered to be consistent and comparable over time. Results are available from 2019-20 to 2022-23. The methodology has been consistent over this time period.

Trends for most farm types in the Farm Business survey are subject to annual sample variations, as a small number of farms join and leave the survey each year. Farms in the sample may also change their characteristics, and might move from being classified as one main farm type to another. More commentary is available in  <a href = "https://www.gov.scot/collections/scottish-farm-business-income-fbi-annual-estimates/" target ="_blank">Scottish farm business income (FBI)</a>.

## Comparability between agriculture sector, subsector and farm-level estimates

Emissions for the agriculture sector and subsectors are for the calendar year and are based on estimates of greenhouse gas (GHG) emissions for Scottish agriculture as reported in the S<a href="https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-2022/" target = "_blank">Scottish Greenhouse Gas Statistics 2022</a>. Agricultural subsectors align more closely with agricultural enterprises and provide a different breakdown to the categories used in the Scottish Greenhouse Gas Statistics.

Nitrogen use estimates for the agriculture sector are available in the <a href ="https://www.gov.scot/publications/scottish-nitrogen-balance-sheet-2021/" target ="_blank">Scottish Nitrogen Balance Sheet</a>.

Average farm estimates do not cover the full agricultural industry and are not directly comparable with national estimates.

Emission and nitrogen use estimates for the average farm are based on commercial sized farms in the Farm Business Survey. Average farm estimates are for the accounting year 2022-23 rather than the 2022 calendar year. Results include all enterprises on farm. Farms are complex businesses with multiple activities that contribute to GHG emissions. For example, cereal farms may have livestock and the scale of this secondary enterprise can vary.

National Scottish estimates use different methodologies that are not directly comparable to the methodology used to estimate farm-level results. Other sources of emissions and nitrogen use data may also use different methodologies. Methodological differences may include:

-   different coefficients to estimate emissions and nitrogen content

-   including or excluding different contributors and use, for example transport of livestock off farm,

-   different methods of accounting for the transfer or life cycle of contributors in the system

-   different industry coverage, for example whether all of meat production is included or results are based on a sub-sample

-   different reporting periods

Users interested in comparing results between different statistics and between different countries should evaluate the relevant methodologies of sources used. Different results may be obtained using different methods.

<br></br>

# Acknowledgements

Thank you to all farmers who participate in the Farm Business Survey. We would like to express our thanks for the commitment and support of all the organisations who contribute to our data gathering exercise. With thanks especially to SAC Consulting who undertake the survey on our behalf. Without your goodwill and support these statistics would not be possible.

<br></br>
